---
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
title:  "Chapter 1-1" 
subtitle:  |
  | Installing R and RStudio
  | The Nature of Econometrics and Econometric Data
author: Jim Bang
description: >
  This tutorial introduces students to the course and demonstrates the basic setup for R and RStudio.
always_allow_html: yes
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

<script language="JavaScript">
$(function() {
   var editor;
   $('.ace_editor').each(function( index ) {
     editor = ace.edit(this);
     editor.setFontSize("16px");
   });
})
</script>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis_setup()
tutorial_options(exercise.eval = FALSE, exercise.checker = gradethis_exercise_checker)
knitr::opts_chunk$set(echo = FALSE, tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Syllabus Information

### General Information

-   Office Hours: MWF 11am-12:30pm; TR 8:30-10am; liberally by appointment: [book me here](https://calendar.google.com/calendar/selfsched?sstoken=UUFLaDRHQXJVeXZHfGRlZmF1bHR8YjU1MzlmMzk2YjBlMWEzMzUzNTUzZDI0MjVhYzBiMTMA)!
-   Research Paper Deadlines:
-   Class Preparation, Participation, and Discussion
    -   The class participant's engagement actively contributes to general student learning -- particularly important with a very small class.
    
### What is Econometrics?

- "How much?"
  - The quantitative analysis of economic data
  - Helps to quantify relationships

### Learning Objectives

To equip students with the tools for analyzing the costs and benefits of trade. 

- Part one of the course will teach the gains from trade; 
- Part two will describe its distributional impact;
- Part three will deal with policy and social issues.

The successful student will be able to: 

1. demonstrate the gains from trade based on comparative advantage; 
2. describe the consequences of trade for income distribution; 
3. analyze the welfare effects of trade intervention; 
4. connect trade theory to other topics 

### Class Resources

1. Blackboard, Inc. 2021. Blackboard. https://blackboard.sau.edu. 
2. R Core Team. 2021. R: A Language and Environment for Statistical Computing. R Foundation for Statistical Computing. 
3. RStudio Team. 2020. RStudio: Integrated Development for R, RStudio, PBC, Boston, MA. 
4. Suranovic, S 2015/2011. International Trade, 2nd Edition (Recommended) or 1st Edition (Minimum Required), FlatWorld, Boston.  

### Class Policies

```{r policies, echo=FALSE}
  question("Which of the following is a class policy?",
    answer("Electronic devices must be turned off at all times"),
    answer("Anyone who attends class and completes all of their own work will earn a C or better", correct = TRUE, message = random_praise()),
    answer("No food or drink allowed in class")
  )
```

### Coursework and Grading

| Assignment Type                                 | Point Total         |
|-------------------------------------------------|---------------------|
| Paper Sections/Drafts (4 @ 50 points each)      | 200 points          |
| Peer Referee Report                             |  50 points          |
| Pilot Analysis Presentation                     |  50 points          |
| Final Draft of Research Paper                   | 250 points          |
| Assessments                                     |  50 points          |
| Participation                                   | 200 points          |
| Final Exam                                      | 200 points          |


<br> <br> 

| Letter Grade     | Point Range               |
|------------------|---------------------------|
| A+/A/A-          | 900-1,000 points          |
| B+/B/B-          | 800-899.9 points          |
| C+/C/C-          | 700-799.9 points          |
| D+/D/D-          | 600-699.9 points          |
| F                | < 600 points              |

## Getting Started with R 

### Why R? 

R helps you talk to a computer and help solve big problems. Employers use R (or other similar languages like Python or Julia) for statistics and data analysis. R is not usually associated with calculus or mathematics, but can do anything that expensive mathematical-analysis programs like Matlab or Mathematica can do - and possibly more! Also, R is free! 

We will use R in this class for two reasons: 

1. As a practical alternative to pencil-and-paper calculus; 
2. To develop basic proficiency with a computing language.

### Installing R

```{r installR, echo=FALSE}
question("Where do you download R?",
    answer("[www.rstudio.com/download](www.rstudio.com/download)"),
    answer("[cloud.r-project.org](http://cloud.r-project.org)", correct = TRUE, message = "You can also download R from [cran.r-project.org](http://cran.r-project.org)"),
    answer("[www.r-project.org](www.r-project.org)", message = "Good try, but not exactly. www.r-project.org doesn't provide a download link, but it does provide a link to one the websites above."),
    answer("[www.r.com](www.r.com)"),
    allow_retry = TRUE
  )
```

### Updating R

```{r updaterR, echo=FALSE}
question("How often should you update R?",
    answer("Everytime you use it", message = "This will be too often unless you use R very rarely!"),
    answer("About once a year", correct = TRUE, "A new version of R is released about once a year. Update sooner if you encounter a bug that you cannot explain."),
    answer("Never", message = "A new version of R is released about once a year. I'll assume that you are using the newest version of R, which will be the fastest version with the fewest unexpected behaviors." ),
    allow_retry = TRUE
  )
```

## RStudio

### What is RStudio?

```{r rStudio, echo=FALSE}
  question("What is the RStudio IDE?",
    answer("An application that makes it easier to use R.", correct = TRUE, message = "RStudio organizes your workspace and makes it easier to write, use, debug, and save R code. I highly recommend it and will rely on the RStudio IDE in several tutorials."),
    answer("An application that let's you use R without writing any code", message = "And thank goodness! Code provides a reproducible record of your work, which is essential for data science."),
    answer("A spreadsheet program like Microsoft Excel."),
    answer("Another name for R", message = "R and RStudio are two separate things. R is a language, like English. Think of RStudio as a program that helps you use the language, kind of like how a word processing program helps you write in English."),
    allow_retry = TRUE
  )
```

### Why RStudio?

RStudio is an Integrated Development Environment (IDE) for R. What does that mean? Well, if you think of R as a language, which it is, you can think of RStudio as a program that helps you write and work in the language. RStudio makes programming in R much easier and I suggest that you use it!

### Where do I get RStudio? 

```{r installRStudio, echo=FALSE}
question("Where do you download RStudio?",
    answer("[www.rstudio.com/download](www.rstudio.com/download)", correct = TRUE, message = "For these tutorials, download and install the RStudio Desktop Open Source License."),
    answer("[cloud.r-project.org](http://cloud.r-project.org)", message = "This is where you download R, not RStudio."),
    answer("[www.r-project.org](www.r-project.org)"),
    answer("[cran.rstudio.org](http://cran.rstudio.org)"),
    allow_retry = TRUE
  )
```

### Do I still need R?

```{r installR_RStudio, echo=FALSE}
  question("Do you need to install R if you already have RStudio?",
    answer("Yes.", correct = TRUE),
    answer("No.", message = "R does not come with RStudio; you need to install R separately.")
  )
```

## R Packages

### How do I install packages?

```{r install_packages, echo = FALSE}
  question("What command do you use to install packages?",
    answer("`library()`", message = "We will see what `library()` does later."),
    answer("`install.packages()`", correct = TRUE),
    answer("`install_packages()`"),
    answer("There is no command. You must visit [cran.r-project.org](http://cran.r-project.org) and download packages manually.", message = "R makes it easy to download packages. You connect to the internet and then run one of the commands above."),
    allow_retry = TRUE
  )
```

### How often do I install a package?

```{r reinstall_packages, echo = FALSE}
  question("How often do you need to install a package on your computer?",
    answer("Every time you restart R"),
    answer("Every time you restart your computer"),
    answer("Once.", correct = TRUE, message = "Afterwards, R can find it on your hard drive and load it using 'library()' or 'require()' as needed."),
    answer("Never, as long as you are connected to the internet.", message = "This could be true if you are using R over a cloud service. However, if you are using R locally on your own computer, you will need to install each package that you use on your computer."),
    allow_retry = TRUE
  )
```

## Economics and the Scientific Method

1.  Ask a Question
2.  Form a Hypothesis
3.  Review Existing Research
4.  Collect Data
5.  Analyze the Data
  - Test *Causal* Hypotheses 
  - Forecast/Predict
6.  Form a Conclusion

### Example 1: Effects of Training on Productivity

$$wage=f(educ,exper,training; u) \text{, where}$$
- $wage$ is the worker's hourly wage,
- $educ$ is a measure of schooling, 
- $exper$ is a measure of workforce experience,  
- $training$ is a measure of time spent training, and
- $u$ is random noise.

Assuming we can approximate $f(\cdot)$ with a linear function, we would 

$$wage=\beta_0 + \beta_1educ + \beta_2exper + \beta_3training + u$$

An econometric model consists of a systematic part – $f(*)$ – and a random error – $u$

### Example 2: Effects of Sentencing on Crime

$$crime=f(crimewage, marketwage, otherincome, p_{caught}, p_{convicted|caught}, \hat{sentence}, age; u) \text{, where}$$ 

- $crime$ is the number of hours spent in illegal activities,
- $crimewage$ is the "wage" per hour for illegal work,
- $marketwage$ is the wage for legal employment, 
- $otherincome$ is income from other sources, 
- $p_{caught}$ is the probability of getting caught, 
- $p_{convicted|caught}$ is the probability of conviction if caught, 
- $sentence$ is the expected sentence if convicted, and 
- $age$ is the person's age.

Some variables might not be observed (crime wage); others might only be observe by proxy or as an average ($p_{caught}$, $p_{convicted|caught}$, $sentence$)

$$crime=\beta_0 + \beta_1marketwage + \beta_2otherinc + \beta_3freqarr + \beta_4freqconv + \beta_5avesen + \beta_6age + u$$

## Data: Sources and Types
-   Experimental versus Observational Data
-   Cross-section
-   Time-series data
-   Pooled cross sections
-   Panel/longitudinal data

## Causality

How could we determine whether spending on education *causes* better performance? 

Hold all else equal: *ceteris paribus*

### Problems: 

#### Confounding (omitted) variables

  Solution: Add those variables to the model.

#### Non-random sample selection

  Solution: Tobit (Chapter 17; Tobin, 1958)
            Heckit (Chapter 17; Heckman, 1976)

#### Non-random treatment assignment

  Solution: Instrumental Variables (Chapter 15)
            Matching treatment and controls (Rosenbaum & Rubin, 1983); 
            Covariate balancing (Horvitz & Thompson, 1952; Robins et al., 2000)
