---
title: "Chapter 14"
subtitle: "Advanced Panel Data Methods"
description: >
  This tutorial practices fixed effects and random effects estimators.
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = TRUE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE, fig.height= 9, fig.width=8)
```

## Panel Data Practice

Replicate Table 14.2 from the book, which estimates the following using the $wagepan$ dataset: 
$$log(wage_{it}) = \beta_0 + \beta_1Education_{it} + \beta_2Black_i + \beta_3Hispanic_i + \beta_4Experience_{it} + $$
$$\beta_5Experience_{it}^2 + \beta_6Married_{it} + \beta_7Union_{it} + \delta_t + a_i + u_{it}$$ 

Estimate pooled OLS, random effects, and fixed effects, as well as correlated random effects.

Note that if we are interested in the race coefficient, fixed effects cannot help us; but if we believe that fixed effects is the appropriate model we can use correlated random effects to eliminate potential bias in and from other variables. 

```{r plmPractice, exercise = TRUE}

```

```{r plmPractice-hint}
wagepan.p <- pdata.frame(df, index)
wagepan.lm <- plm(formula, data, model)
wagepan.fe <- plm(formula, data, model)
wagepan.re <- plm(formula, data, model)
stargazer(list, type)
```

```{r plmPractice-solution, exercise.reveal_solution = FALSE}
wagepan.p <- pdata.frame(wagepan, index=c("nr","year") )
wagepan.lm <- plm(lwage~educ+black+hisp+exper+I(exper^2)+married+union+factor(year), 
                  data=wagepan.p, model="pooling")
wagepan.re <- plm(lwage~educ+black+hisp+exper+I(exper^2)+married+union+factor(year), 
                  data=wagepan.p, model="random")
wagepan.fe <- plm(lwage~educ+black+hisp+exper+I(exper^2)+married+union+factor(year), 
                  data=wagepan.p, model="within")
wagepan.cre <- plm(lwage~educ+black+hisp+exper+I(exper^2)+married+union+factor(year)
                   +Between(exper)+Between(I(exper^2))+Between(married)+Between(union), 
                  data=wagepan.p, model="random")

stargazer(wagepan.lm, wagepan.re, wagepan.fe, wagepan.cre, type = 'text')
```

```{r plmPractice-check}
grade_code()
```

