---
title: "Chapter 8-2: Estimation with Heteroskedasticity"
author: Jim Bang
description: "Consequences of heterodkedasticity and testing for heteroskedasticity"
output: 
  learnr::tutorial: 
    progressive: true
runtime: shiny_prerendered
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

<script language="JavaScript">
$(function() {
   var editor;
   $('.ace_editor').each(function( index ) {
     editor = ace.edit(this);
     editor.setFontSize("16px");
   });
})
</script>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(lmtest)
library(sandwich)
library(car)
library(stargazer)
tutorial_options(exercise.timelimit = 60, exercise.reveal_solution = TRUE)
knitr::opts_chunk$set(echo = FALSE)
```

## Robust Standard Errors, Revisited

Recall that $Var(\hat\beta) = (X'X)^{-1}X'X(X'X)^{-1}\sigma_\hat{u}^2$, <br>
which we can rewrite as $(X'X)^{-1}X'(\sigma_\hat{u}^2I)X(X'X)^{-1}$.

Some people make an analogy for this form to a "sandwich" with an outer "bread" consisting of the invers of the sums of squares and cross-products, $(X'X)^-1$, and "meat" consisting of $X'\hat\sigma_\hat{u}^2IX$.

We can rewrite the "meat" as $\Omega$, and $Var(\hat\beta)$ as $(X'X)^{-1}\Omega(X'X)^{-1}$ Fixes for heteroskedasticity focus on the structure of the "meat."

There are many types of robust, or *heteroskedasticity-consistent* - standard errors. 

### Basic, `HC1`-type, Heteroskdasticity-Robust $Var(\hat\beta)$



Estimate the log wage equation with (`wage.lm14`) and without (from the book. 

```{r robustHC1, exercise = TRUE}

```

```{r robustHC1-solution}
wage.lm14 <-
  lm(
    log(wage) ~ I(married * (1 - female)) + I(married * female) + 
      I((1 - married) * female) + educ + exper + I(exper ^ 2) + tenure + 
      I(tenure ^ 2), data = wage1
  )
wage.lm14r <-
  coeftest(wage.lm14, vcov = vcovHC(wage.lm14, type = "HC1"))
stargazer(wage.lm14, wage.lm14r, type = 'text', digits = 4)
```

```{r robustHC1-check}
grade_code()
```

### The `sandwich::vcovHC` Default: $HC3$



## Weighted Least Squares


