---
title: "Chapter 6"
subtitle: "Multiple Regression Analysis - "
description: "This tutorial practices interactions and comparing specification."
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = FALSE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```

## Square Footage, Bedrooms, and Home Prices 

Estimate the following models using the `hprice1` dataset.

1. $Price = \beta_0 + \beta_1Lot\text{ }Size + \beta_2Square\text{ }Feet + \beta_3Bedrooms + u$
2. $Price = \beta_0 + \beta_1Lot\text{ }Size + \beta_2Square\text{ }Feet + \beta_3Bedrooms + \beta4Square\text{ }Feet \cdot Bedrooms + u$
3. $Price = \beta_0 + \beta_1Lot\text{ }Size + \beta_2Square\text{ }Feet + \beta_3Bedrooms + \beta4(Square\text{ }Feet - \mu_{Sq.\text{ }Ft.}) \cdot (Bedrooms - \mu_{Bedrooms}) + u$

Present all three sets of estimates in the same **text** table using `stargazer.`

```{r housePricePractice, exercise=TRUE, exercise.reveal_solution = FALSE}

```

```{r housePricePractice-solution}
library(stargazer)
hprice1 <- wooldridge::hprice1
hprice.lm1 <- lm(price ~ lotsize + sqrft + bdrms, data = hprice1)
hprice.lm2 <- lm(price ~ lotsize + sqrft*bdrms, data = hprice1)
hprice.lm3 <- lm(
  price ~ lotsize + sqrft + bdrms + I(scale(sqrft, T, F)*scale(bdrms, T, F)),
  data = hprice1)
stargazer(hprice.lm1, hprice.lm2, hprice.lm3, type = 'text')
```

```{r housePricePractice-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Comparing Nested Specifications

Add the squares of square footage and number of bedrooms using the centered interactions from the third specification in the previous question.  
Calculate and store the AICs for all three unique models (no interaction and no squares, interaction but no squares, interaction with squares).  Return a `c()` vector including the AIC for all three specifications (only linear terms, linear with scaled interaction, and linear with interactions and quadratics).

```{r AIC, exercise = TRUE}

```

```{r, AIC-setup}
library(stargazer)
hprice1 <- wooldridge::hprice1
hprice1 <- wooldridge::hprice1
hprice.lm1 <- lm(price ~ lotsize + sqrft + bdrms, data = hprice1)
hprice.lm2 <- lm(price ~ lotsize + sqrft*bdrms, data = hprice1)
hprice.lm3 <- lm(
  price ~ lotsize + sqrft + bdrms + I(scale(sqrft, T, F)*scale(bdrms, T, F)),
  data = hprice1)
```

```{r, AIC-hint}
# The exercise setup includes the `lm()` objects from the previous example as `hprice.lm1`, `hprice.lm2`, and `hprice.lm3`. 
```

```{r, AIC-solution}
hprice.lm4 <- lm(
  price ~ lotsize + sqrft + bdrms + I(scale(sqrft, T, F)*scale(bdrms, T, F)) +  
  I(sqrft^2) + I(bdrms^2), data = hprice1)
c(AIC(hprice.lm1), AIC(hprice.lm3), AIC(hprice.lm4))
```

```{r, AIC-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Adding Diagnostic Stats to `stargazer()` Output

You may specify specific statistics to add or keep for your table output in `stargazer()` using the `keep.stat` (if `stargazer` supports calcculating a specific internally for your model type) or `add.lines`.

`stargazer` does support calculating AICs, but for some reason not for simplie `lm()` classes. Add the AIC to the table for each of the three models you calculated AIC for in the previous exercise. 

```{r stargazer, exercise = TRUE}

```

```{r, stargazer-setup}
library(stargazer)
hprice1 <- wooldridge::hprice1
hprice.lm1 <- lm(price ~ lotsize + sqrft + bdrms, data = hprice1)
hprice.lm3 <- lm(
  price ~ lotsize + sqrft + bdrms + I(scale(sqrft, T, F)*scale(bdrms, T, F)),
  data = hprice1)
hprice.lm4 <- lm(
  price ~ lotsize + sqrft + bdrms + I(scale(sqrft, T, F)*scale(bdrms, T, F)) +  
  I(sqrft^2) + I(bdrms^2), data = hprice1)
hprice.aic1 <- AIC(hprice.lm1) 
hprice.aic3 <- AIC(hprice.lm3)
hprice.aic4 <- AIC(hprice.lm4)
```

```{r, stargazer-hint-1}
# The exercise setup includes the `AIC()` statistics from the previous example as `hprice.aic1`, `hprice.aic3`, and `hprice.aic4`. 
```

```{r, stargazer-hint-2}
# `add.lines` takes a `list()` of `c()` vectors, where each vector in the list specifies the values for each column of the line it adds, including row label. 
# `add.lines = list(c("Label for first line added", column1value, column2value, ...))
```

```{r, stargazer-solution}
stargazer(
  hprice.lm1, hprice.lm3, hprice.lm4, type = 'text', 
  add.lines = list(
    c("AIC", round(hprice.aic1, 3), round(hprice.aic3, 3), round(hprice.aic4, 3))
    )
  )
```

```{r, stargazer-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```
