---
output: slidy_presentation
runtime: shiny_prerendered
title: "Chapter 13 \n\n Pooled Cross Sections and \n\n Difference-in-Differences Estimation"
author: "James Bang"
description: >
  This tutorial introduces simple panel methods.
---

```{r setup, include=FALSE}
library(learnr)
library(wooldridge)
library(stargazer)
wage1 <- wooldridge::wage1
wage.lm7 <- lm(wage ~ educ + exper + I(exper^2) + tenure + profocc + clerocc + servocc, data = wage1)
wage.lm9 <- lm(wage ~ educ + exper + tenure + female, data = wage1)
wage.lm11 <- lm(wage ~ educ + exper + tenure + female*married, data = wage1)
wage.lm12 <- lm(log(wage) ~ educ + exper + tenure + female*married, data = wage1)
wage.lm13 <- lm(wage ~ educ + exper + tenure + female*educ, data = wage1)
knitr::opts_chunk$set(echo = FALSE)
```

## Pooled Cross Sections vs. Panel (Longitudinal) Data

Pooled Cross Sections - Data obtained by conducting a cross-sectional survey multiple times over different time periods with independent random samples each period. 

  - Current Population Survey
  - World Values Survey (Individual Responses)
  - General Social Survey
  
Panel (Longitudinal) Data - Data obtained by tracking the same (randomly-sampled) individuals over multiple time periods. 

  - Panel Study of Income Dynamics
  - World Development Indicators
  - World Values Survey (Country Averages)
  - Various Business Sources at:
  
  https://maryannfeldman.web.unc.edu/data-sources/longitudinal-databases/

## Single Dummy Independent Variable



```{r gender, exercise = TRUE}

```

```{r gender-hint}

```

```{r gender-solution}

```

```{r gender-check}
grade_code()
```

## Dummy Variables for Multiple Categories



```{r occupation, exercise = TRUE}

```

```{r occupation-hint}

```

```{r occupation-solution}

```

```{r occupation-check}
grade_code()
```

## Interactions among Dummy Variables



```{r marriage, exercise = TRUE}

```

```{r marriage-hint}

```

```{r marriage-solution}

```

```{r marriage-check}
grade_code()
```

