---
output: pdf_document
classoption: landscape
title: "Appendix B \n\n Fundamentals of Probability"
description: "This tutorial reviews some basic concepts about relationships between two variables."
---

```{r setup, include=FALSE}
ceosal1 <- wooldridge::ceosal1
affairs <- wooldridge::affairs
attach(ceosal1) 
E_sal.roe <- mean(salary*roe)
E_sal.E_roe <- mean(salary)*mean(roe)
affairs$haskids <- factor(affairs$kids, labels = c("no","yes"))
affairs$marriage <- factor(affairs$ratemarr, labels = c("very unhappy","unhappy","average","happy", "very happy"))
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Joint Distributions, Conditional Distributions, and Independence

### Joint Distributions

-   Discrete Case: $f_{X,Y}(x,y) = P(X = x, Y = y)$
-   Continuous Case: $f_{X,Y}(x,y)$ is its own function.

For *independent* random variables the joint distribution can be separated as the product of the marginal distributions.

-   Discrete Case: $f_{X,Y}(x,y) = P(X = x) \cdot P(Y = y)$
-   Continuous Case: $f_{X,Y}(x,y) = f_X(x) \cdot f_Y(y)$

### Conditional Distributions

-   Discrete Case: $f_{Y|X}(y|x) = P(X = x, Y = y)/P_x(X=x)$
-   Continuous Case: $f_{Y|X}(y|x) = f_{X,Y}(x, y)/f_X(x)$

For *independent* random variables, knowing information about $X$ does not change the probability of $Y$.  

-   Discrete Case: $P_{Y|X}(y|x) = P(Y = y) \cdot P(Y = y)/P(X = x) = P(Y = y)$
-   Continuous Case: $f_{Y|X}(y|x) = f_Y(y) \cdot f_X(x)/f_X(x)$

\newpage 
### *Estimated* Joint Probabilities of Discrete Variables: Crosstabulations

Sample proportions, or relative frequencies, estimate the population proportions (probabilities) of an outcome.

Re-create the crosstabulation of marital happiness and whether a couple has kids from the affairs dataset showing only the *proportion* (not percentage) of each joint outcome.

```{r joint-solution}
prop.table(table(affairs$marriage,affairs$haskids))
```

\newpage 
### *Estimated* Conditional Probabilities of Discrete Variables: Row Margins

Conditional probabilities only consider the probablities of $Y$ *given* a specific known outcome on $X$ (or vice-versa). The `margins` option in `prop.table` conditions the relative frequencies on knowing the outcome of the row (`margins = 1`) or column (`margins = 2`). 

Re-create the crosstabulation of marital happiness and whether a couple has kids from the affairs dataset showing only the *proportion* (not percentage) of each *conditional* outcome.

```{r conditional-solution}
prop.table(table(affairs$marriage,affairs$haskids), margin = 1)
prop.table(table(affairs$marriage,affairs$haskids), margin = 2)
```

\newpage
## Covariance and Correlation

Covariance and correlation are (related) measures of the linear association between two variables. 

### Covariance

$$Cov(X,Y) = \sigma_{XY} = E[(X - \mu_X)(Y - \mu_Y)]$$

Distributing within the expected value, covariance "simplifies" to: 

$$\sigma_{XY} = E(XY) - E(X)E(Y)$$

Variance is actually a special case of covariance, where $X = Y$. 

$$V(X) = Cov(X,X) = \sigma_{XX} = E[(X - \mu_X)(X - \mu_X)]$$

### Properties of Covariance

1. Adding a constant to either variable changes nothing: for any constants, $a_1$ and $a_2$, $Cov(a_1 + X, a_2 + Y) = Cov(X,Y)$
2. Multiplication factors out: for any constants, $a_1$, $a_2$, $b_1$, and $b_2$, $Cov(a_1 + b_1X, a_2 + b_2Y) = b_1b_2Cov(X,Y)$
3. Linear functions: for any constants, $b_1$ and $b_2$, $Cov(b_1X,Y) = b_1Cov(X,Y)$, $Cov(X,b_2Y) = b_2Cov(X,Y)$, and  $Cov(b_1X,b_2Y) = b_1b_2Cov(X,Y)$
4. If $X$ and $Y$ are independent, $Cov(X,Y) = 0)$. 
5. Covariance is bounded: $|Cov(X,Y)| \le \sigma_X \cdot \sigma_Y$

### Covariance Example

Using the ceosal1 data, calculate: 

- $E(Salary \cdot ROE)$ (name this "E_sal.roe")
- $E(Salary) \cdot E(ROE)$ (name this "E_sal.E_roe")
- $Cov(Salary,ROE) = E(Salary \cdot ROE) - E(Salary) \cdot E(ROE)$

```{r Exy-solution}
attach(ceosal1) 
E_sal.roe <- mean(salary*roe)
E_sal.E_roe <- mean(salary)*mean(roe)
E_sal.roe - E_sal.E_roe
```

### Correlation Coefficient

The correlation coefficient is a standardized measure of association. 

$$Corr(X,Y) = \rho_{XY} = \frac{Cov(X,Y)}{\sqrt{V(X) \cdot V(Y)}} = \frac{\sigma_{XY}}{\sigma_X \cdot \sigma_Y}$$

### Correlation Example
```{r cor-solution}
(E_sal.roe - E_sal.E_roe)/(sd(salary)*sd(roe))
```

### Using `cov` and `cor` Functions

Calculate the covariance and correlation for salary and roe using the "cov" and "cor" functions.

```{r covcor2-solution}
cov(ceosal1$salary, ceosal1$roe)
cor(ceosal1$salary, ceosal1$roe)
```

### Correlation and Independence

  - What is $corr(x, x^2)$?

  - Are $x$ and $x^2$ independent?",

\newpage 
## Conditional Expectation

For a discrete random variable, $Y \in \{y_1, y_2, ..., y_m \}$, the conditional expected value of $Y$ is given by 

$$E(Y|X) = \sum_{j=1}^m y_j f_{Y|X}(y_j|x)$$
For a continuous random variable, 

$$E(Y|X) = \int_{y \in \Omega_Y} y_j f_{Y|X}(y|x)dy$$

### Properties of Conditional Expectation

1. For any function, $c(X)$, $E[c(X)|X] = c(X)$
2. For *any* functions, $a(X)$ and $b(X)$, $E[a(X)Y + b(X) | X] = a(X)E(Y|X) + b(X)$ For example, $E(XY + 2x^2 | X) = XE(Y|X) + 2X^2$
3. If *X* and *Y* are independent, $E(Y|X) = E(Y)$
4. If $E(Y|X) = E(Y)$, then $Cov(X,Y) = 0$ *and* for *any* function, $f(X)$, $Cov(f(X), Y) = 0$
5. If $E(Y^2) < \infty$ and for some function $g$, $E(g(X)^2) < \infty$ then: 
  a. The conditional mean is better (based on expected squared prediciton error) than any other function of $X$ for predicting $Y$, conditional on $X$: <br>
  $E[(Y - \mu_X)^2|X] \le E\{[Y - g(X)]^2|X\}$ for any function $g(X) \ne \mu_X$
  b. The conditional mean minimizes the unconditional expected squared prediction error: <br>
  $E[(Y - \mu_X)^2] \le E\{[Y - g(X)]^2\}$ for any function $g(X) \ne \mu_X$
6. Law of Iterated Expectations 
  a. The expected value over $X$ of the expected value of $Y$ *conditional on* $X$ over all values of $X$ equals the unconditional mean of $Y$: $E[E(Y|X)] = E(Y)$.
  b. Given random variables $X$ and $Z$, we can find the expected value of $Y$ conditional on $X$ in two steps. 
    i. Find $E(Y|X,Z)$
    ii. Take the expected value of the result conditional on $X$ 

$$E(Y|X) = E[E(Y|X,Z)|X]$$

### Conditional Variance

$$V(Y|X = x) = E(Y^2|x) - [E(Y|x)]^2$$

\newpage 
## Normal (and Other) Distributions

### Univariate Normal Distribution

$$ f(x) = \frac{1}{\sigma_x \sqrt{2\pi}}e^{\frac{-(x - \mu_x)^2}{2\sigma_x^2}}$$ 

### Standard Normal Distribution

$$ f(z) = \frac{1}{\sqrt{2\pi}}e^{\frac{-z^2}{2}}$$ 

### Bivariate Normal Distribution (Independent Ranodm Variables): 

$$ f(x,y) = f(x) \cdot f(y) = \frac{1}{2\pi\sigma_x\sigma_y}e^{-\frac{1}{2}([\frac{(x - \mu_x)}{\sigma_x}]^2+[\frac{(y - \mu_y)}{\sigma_y}]^2)} $$ 

### Bivariate Normal Distribution (Correlated Random Variables): 

$$ f(x,y) = \frac{1}{2\pi\sigma_x\sigma_y\sqrt{1-\rho^2}}e^{\frac{-1}{2(1-\rho^2)}[(\frac{x-\mu_x}{\sigma_x})^2-2\rho[(\frac{x-\mu_x}{\sigma_x})(\frac{y-\mu_y}{\sigma_y})]+(\frac{y-\mu_y}{\sigma_y})]^2} $$

### Multivariate Normal Distribution (Matrix Form): 

$$ f(\mathbf{x}) = \frac{1}{\sqrt{2\pi|\mathbf{\Sigma}|}}e^{(\mathbf{x}-\mathbf{\mu})'\mathbf{\Sigma}^{-1}(\mathbf{x}-\mathbf{\mu})} $$

### Other Distributions

- Student's *t* Distribution
- Chi-Square Distribution
- *F* Distribution
