---
output: pdf_document
classoption: landscape
title: "Chapter 19 \n\n Conducting Research \n\n Presenting Tables and Graphs"
description: "This tutorial introduces tools for creating attractive tables and graphs."
---

```{r setup, include=FALSE}
library(gtsummary)
library(corrgram)
library(ggplot2)
library(wooldridge)
ceosal1 <- wooldridge::ceosal1
affairs <- wooldridge::affairs
affairs$haskids <- factor(affairs$kids, labels = c("no","yes"))
affairs$marriage <- factor(affairs$ratemarr, labels = c("very unhappy","unhappy","average","happy", "very happy"))
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Doing Economics: A Primer

### Posing a question

- Topic – the beginning of your search, not the end
  - Narrow it down
  - Talk to (other) faculty
  - Research existing literature
  - Find a niche or replicate?
- Assess feasibility
  - Are data sufficient?
  - Are my tools and other resources sufficient? 
- Focus on a testable hypothesis

### Literature Review

- Review the *literature*, not individual contributions
  - Organize papers into different “threads” of the literature
  -Differentiate by type of analysis (theory vs. empirical, causal vs. descriptive), type of result (positive, negative, none), type of data (geographic location, macro vs. micro, time periods), sub-topics…
- State your paper’s contribution
  - Niche 
  - Replication
  - ~~Literature Review~~
  - ~~Meta-Analysis~~
- Databases

### Research Databases

- [Library](http://www.sau.edu/Library/Resources/Find_Articles.html): EBSCO Host, JSTOR, Science Direct
- [Google Scholar](http://scholar.google.com)
  - Pro tip: Select "Endnote" as your bibliography manager in "Search results" and search and add "SAU" to the "Library links" in [settings](https://scholar.google.com/scholar_settings). 

### Structure of the Paper 

- Introduction
- Conceptual Framework
- Data and Methods
- Results and Discussion
- Conclusion

### Data Collection

- Online Sources
  - Formatted Electronic Data (.xls, .csv, .dta, .dat, etc….)
  - PDFs (require some manual entry/editing)
- Merging Data 
  - ID variable (Key)
  - Dealing with mismatches?
- Cleaning Data
  - Dealing with (and reformatting) NA’s
  - Naming, labelling, and storing data (.Rdata files, .Rproj files, and project directories) 

\newpage
## Summarizing Data

### Single Variables

Do the following: 
- Explore the objects in your environment using *ls()*
- Using the *ceosal1* data, separately calculate the mean, median, and standard deviation of the *salary* variable. 
- Calculate the summary statistics for the *salary* variable.

```{r singleSummary}
ls()
mean(ceosal1$salary)
median(ceosal1$salary)
sd(ceosal1$salary)
summary(ceosal1$salary)
```

### Summarizing an Entire Data Frame

- Calculate the summary statistics for all of the variables in *ceosal1*. 

```{r multipleSummary}
summary(ceosal1)
```

\newpage
## Creating Pretty Tables

### Standard Output

Create summary tables of *ceosal1* using the *tbl_summary* function in the *gtsummary* package.

```{r gtmedian}
tbl_summary(ceosal1)
```

### Means and Standard Deviations

Input the request from the feedback below.

```{r gtsummary}
tbl_summary(ceosal1, statistic = list(all_continuous() ~ "{mean} ({sd})"), digits = list(all_continuous() ~ c(2,2)))
```

\newpage
## Relationships between Variables

### Correlation Matrix

Calculate the following:

- The correlation of salary with ROE;
- The correlation matrix for all numeric variables.

```{r correlation}
cor(ceosal1$salary, ceosal1$roe)
cor(ceosal1)
```

### Pretty Correlation Matrix

Use the *corrgram* function (part of the *corrgram* package) to create a correlogram.

```{r corrheatmap}
corrgram(ceosal1)
```

### Pretty *Numerical* Correlation Matrix

Input the suggestion from the feedback below. 

```{r corrgram}
corrgram(ceosal1, panel = panel.cor)
```

\newpage
## Factor Variables

### Crosstabulation Tables

Using the *affairs* data, do the following: 

- Generate the factor variables *haskids* and *marriage* for kids and ratemarr labels with labels *no*/*yes* and *very unhappy*/*unhappy*/*average*/*happy*/*very happy*, respectively;
- Create a table that displays the proportions of each outcome of marriage happiness rating; 
- Create a table that displays the proportions of each outcome of marriage happiness rating *and having kids*.

```{r crosstab}
affairs$haskids <- factor(affairs$kids, labels = c("no","yes"))
affairs$marriage <- factor(affairs$ratemarr, labels = c("very unhappy","unhappy","average","happy", "very happy"))
prop.table(table(affairs$marriage))
prop.table(table(affairs$marriage,affairs$haskids))
```

### Pretty Crosstab Tables

- Replicate the previous tables using "tbl_cross" (using percent to generate percentages instead of proportions) to generate a more attractive layout that you can save as html. 
- Combine counts and percentages in one table with each cell displaying percentages in parentheses next to its count.

```{r tbl_cross}
tbl_cross(affairs, row = marriage, col = haskids)
tbl_cross(affairs, row = marriage, col = haskids, statistic = "{p}%")
tbl_cross(affairs, row = marriage, col = haskids, percent = 'cell')
```

\newpage
## Graphs 

### Base R Graphics 

Plot the following: 

- A histogram of CEO salary *relative* frequencies using the base-graphics hist() function;
- A scatterplot of CEO salaries with ROE using the plot() function.

Make sure your plots have descriptive (English) titles: *Histogram of Salary*, *Salary versus ROE*, *Salary*, and *ROE*.

```{r basePlots}
hist(ceosal1$salary, main = "Histogram of Salary", xlab = "Salary", freq = FALSE)
plot(ceosal1$salary, ceosal1$roe, main = "Salary versus ROE", xlab = "Salary", ylab = "ROE")
```

### Grammar of Graphics (ggplot2)

Some people prefer base graphics. Some prefer gg. Depending on what packages you use, you should know a little of each. You can call the ggplot() function with no arguments (but add them in layers) or inline. Either way, the main arguments you need to specify include (1) data and (2) aesthetics (aes). 

Try replicating the previous plots using ggplot geom_hist() and geom_point syntax. 

```{r ggplot}
ggplot() +
  geom_histogram(data = ceosal1, mapping = aes(salary))
ggplot() + 
  geom_point(data = ceosal1, mapping = aes(salary, roe))
```

### Options in `ggplot`

Implement the suggestion in the feedback. 

```{r ggrelfreq}
ggplot() +
  geom_histogram(data = ceosal1, mapping = aes(x = salary, y = stat(count)/sum(count)))
```
