---
output: slidy_presentation
runtime: shiny_prerendered
title: "Chapter 13 \n\n Pooled Cross Sections and \n\n Difference-in-Differences Estimation"
author: "James Bang"
description: > 
  This tutorial provides practice excercies on simple panel data methods.
---

```{r setup, include=FALSE}
library(learnr)
library(wooldridge)
library(plm)
library(gtsummary)
library(stargazer)
library(gradethis)
gradethis_setup()
traffic1 <- wooldridge::traffic1
traffic1.long <- reshape(traffic1[,1:9], varying = c('admn90', 'admn85', 'open90', 'open85', 'dthrte90', 'dthrte85', 'speed90', 'speed85'), timevar = 'year', idvar = 'state', direction = 'long', sep = '')
traffic1.long$time <- traffic1.long$year/5
traffic1.lm1 <- lm(dthrte ~ open + admn + factor(year) + factor(state), data = traffic1.long)
traffic1.pdata <- pdata.frame(traffic1.long, index = c('state', 'time'))
knitr::opts_chunk$set(echo = FALSE)
```

## Two-Period Panal Data Analysis Using First Differences

Create a panel from the state id, administrative per se law dummy, open-container law dummy, and year id variables in the traffic1 data, and name it traffic1.pdata
  1. Use $reshape$ to convert the $admn$, $open$, $speed$, and $dthrte$ variables (rows 1 through 9, including the state id variable) from wide (separate columns for 1985 and 1990) to long (stacked rows) format ($traffic1.long$). 
  2. Use $pdata.frame$ to convert the structure from pooled cross sections to panel format ($i = state$, $t = year$: $traffic1.pdata$).

```{r reshape, exercise = TRUE}

```

```{r reshape-hint}
traffic1.long <- reshape(data[rows, columns], varying = variable.names, timevar = newname, idvar = oldname, direction, sep)
traffic1.pdata <- pdata.frame(traffic1.pdata, index = c('state', 'year'))
```

```{r reshape-solution, exercise.reveal_solution = FALSE}
traffic1.long <- reshape(traffic1[,1:9], varying = c('admn90', 'admn85', 'open90', 'open85', 'dthrte90', 'dthrte85', 'speed90', 'speed85'), timevar = 'year', idvar = 'state', direction = 'long', sep = '')
traffic1.long$time <- traffic1.long$year/5
traffic1.pdata <- pdata.frame(traffic1.long, index = c('state', 'time'))
```

```{r reshape-check}
grade_code()
```

Estimate the following: 

  1. $DeathRate_{it} = \beta_0 + \beta_1OpenContainerLaw_{it} + \beta_2Administrative_{it} + \delta_t + a_i + u_{it}$ using the $lm$ function ($death.lm1$).
  2. $\Delta DeathRate_{it} = \beta_0 + \beta_1\Delta beta_1OpenContainerLaw_{it} + \beta_2Administrative_{it} + \Delta u_{it}$ using the $lm$ function ($death.lm2$).
  3. $CrimeRate_{it} = \beta_0 + \beta_1OpenContainerLaw_{it} + \beta_2Administrative_{it} + u_{it}$ using the first-differences ($'fd'$) model option of the $plm(formula, pdata, model)$ function ($death.plm1$).
  
Summarize all three models in text format using the $stargazer$ function. Omit the year and id dummy variables from the first specification. 

```{r death, exercise = TRUE}

```

```{r death-hint}

```

```{r death-solution, exercise.reveal_solution = FALSE}
death.lm1 <- lm(dthrte ~ open + admn + factor(year) + factor(state), data = traffic1.long)
traffic1.pdata$ddthrte <- diff(traffic1.pdata$dthrte)
traffic1.pdata$dadmn <- diff(traffic1.pdata$admn)
traffic1.pdata$dopen <- diff(traffic1.pdata$open)
death.lm2 <- lm(ddthrte ~ dopen + dadmn, data = traffic1.pdata)
death.plm1 <- plm(dthrte ~ open + admn, data = traffic1.pdata, model = 'fd')
stargazer(death.lm1, death.lm2, death.plm1, type = 'text', omit = c('year', 'state'))
```

```{r death-check}
grade_code()
```

## Differencing More than Two Time Periods

Estimate the following specification using the $mathpnl$ dataset: 

$$MathPassRate_{it} = \beta_0 + \beta_1ln(RealExpPerStudent_{it}) + \beta_2ln(Enrollment_{it}) + \beta_3LunchRate_{it} + \delta_t + \alpha_i + u_{it}$$
Then, create a $pdata.frame$ object ($mathpnl.pdata$) to estimate the analogous model using first differences and compare the results using a stargazer text table. 

```{r math, exercise = TRUE}

```

```{r math-hint}
```

```{r math-solution, exercise.reveal_solution = FALSE}
math.lm1 <- lm(math4 ~ log(rexpp) + log(enrol) + lunch + factor(year) + factor(distid), data = mathpnl)
mathpnl.pdata <- pdata.frame(mathpnl, index = c('distid', 'year'))
math.plm1 <- plm(math4 ~ log(rexpp) + log(enrol) + lunch, data = mathpnl, model = 'fd')
stargazer(math.lm1, math.plm1, type = 'text', omit = c('year', 'distid'))
```

```{r fd-check}
grade_code("These aren't necessarily the same with T > 2!")
```
