---
title: "Chapter 1-1 Practice"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simple Calculations

Write the R code required to perform the following operations:

- $21 + 21$
- $5(4-1)^2-3$
- $\sqrt{1764}$

```{r R_as_a_calculator-solution, results='hold'}
21+21
5*(4-1)^2-3
sqrt(1764)
```

## Objects

R is an object-oriented language. The "assignment" operator, "<-" (keyboard shortcut: "Alt" + "-") computes the operation to the right and gives it name in your *environment* (see top-right pane in RStudio). Name the calculations from the previous exercise as indicated below.

- $21 + 21$: answer
- $5(4-1)^2-3$: life
- $\sqrt{1764}$: universe
- $ln(e^{42})$: everything

Display the objects in your environment using the `ls()` function. 

```{r objects-solution}
answer <- 21+21
life <- 5*(4-1)^2-3
universe <- sqrt(1764)
everything <- log(exp(42))
ls()
```

## Functions

If you can define an object you can write (and evaluate) your own function! A function can be a simple mathematical function, or it can be a long script that outputs several different values. Define the function, `quadratic`, as a function of *variable* `x` and *parameters* `a`, `b`, and `c` as 
$$f(x) = ax^2 + bx + c.$$ 
Then, evaluate $f(x) = 8x^2 - 6x - 12$ at $x = 3$

```{r functions-solution}
quadratic <- function(x, a, b, c) {
  a*x^2 + b*x + c
}
quadratic(3, 8, -6, -12)
```

For the purposes of getting computational solutions, there is not really a difference between a parameter and a variable. Economically, we consider parameters (a, b, c, and d) to be *exogenous* - determined by nature (like preference parameters) or markets (like prices or income) - whereas we consider variables (x) to be *endogenous* - chosen by households (like quantities consumed) or firms (like amounts of a resource hired). 

## Vectors

- Create the vector $a = \begin{pmatrix} 1 \\ 2 \\ 3 \\ 4 \\ 5 \\ 6 \end{pmatrix}$;
- Assign the value $a + 1$ to the object b;
- Assign the value $a + b$ to the object c;
- Assign the value $b + c$ to the object d;
- Calculate $\sqrt{d}$;

```{r Vectors-solution}
a <- c(1,2,3,4,5,6)
b <- a+1
c <- a+b
d <- b+c
sqrt(d)
```

## Matrices

- Create a matrix, A that splits the values of a across two columns;
- Create a matrix, B that duplicates the columns of A twice. 
- Extract the third column of B as a matrix named "b."

```{r Matrices-solution}
A <- matrix(a, ncol = 2)
B <- cbind(A, A);
b <- B[, 3]
```

## Viewing

Display the first *five* rows of the mtcars dataset. Then, View the complete data frame in a new window.

```{r View-solution}
head(mtcars, n = 5)
View(mtcars)
```
