---
title: "Chapter 13 \n\n Pooling Cross Sections across Time: \n\n Simple Panel Data Methods"
description: >
  This tutorial provides practice with pooled cross sections and difference-in-difference methods.
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = TRUE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE, fig.height= 9, fig.width=8)
```

## Pooled Cross Sections: Fertility over Time

Estimate the determinants of women's fertility using the $fertil1$ data, controlling for age, age-squared, race (black vs. non-black), region, and geographic type (city, farm, rural, town, small city). Name this $fertil.lm1$.

Next, estimate the "non parametrice time trend" (year dummies) for fertililty over time. Name This $fertil.lm2$.

Summarize both regressions using a "stargazer" text-table.

```{r fertility, exercise = TRUE}

```

```{r fertility-hint}
fertil.lm1 <- lm(formula, data)
fertil.lm2 <- lm(formula, data)
stargazer(lm-object, ..., type)
```

```{r fertility-solution, exercise.reveal_solution = FALSE}
fertil.lm1 <- lm(kids ~ educ + age + I(age^2) + black + region + geog, data = fertil1)
fertil.lm2 <- lm(kids ~ educ + age + I(age^2) + black + region + geog + factor(year), data = fertil1)
stargazer(fertil.lm1, fertil.lm2, type = "text")
```

```{r fertility-check}
grade_code()
```

## Difference-in-Difference: Worker Compensation and Weeks Out-of-Work

Estimate the effect on worker's compensation duration from a change in worker's compensation that raised the earnings cap on benefits. The simple DiD model for this would look like:

$$Duration_i = \beta_0 + \delta_0After Change + beta_1HighEarner_i + \delta_1AfterChanges*HighEarner_i +  u_{it}$$ Name this $workcomp.lm1$.

Next, estimate the same model while also controlling for gender ($male$) and type of injury ($injtype$). Name this specification $workcomp.lm2$.

Summarize both regressions using a "stargazer" text-table.

```{r workcomp, exercise = TRUE}

```

```{r workcomp-hint}
workcomp.lm1 <- lm(formula, data)
workcomp.lm1 <- lm(formula, data)
stargazer(lm-object, ..., type)
```

```{r workcomp-solution}
workcomp.lm1 <- lm(log(durat) ~ afchnge*highearn, data = injury)
workcomp.lm1 <- lm(log(durat) ~ afchnge*highearn + male + injtype, data = injury)
stargazer(fertil.lm1, fertil.lm2, type = "text")
```

```{r workcomp-check}
grade_code()
```
