---
title: "Chapter 15"
subtitle: "Instrumental Variables Methods"
description: >
  This tutorial introduces instrumental variables. 
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = TRUE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE, fig.height= 9, fig.width=8)
```

## Causal Diagrams (Directed Acyclic Graphs, DAGs)

The effect we wish to estimate is: 

$$ln(Wage_i) = \beta_0 + \beta_1Education_i + u_i$$

```{r, include = TRUE, fig.align = 'center'}
ggdag(WageDag1) + theme_dag(legend.position = 'none') + geom_dag_label_repel(aes(label = label))
```

## Causal Diagrams (Directed Acyclic Graphs, DAGs)

The code that creates the simple DAG is:

```{r, echo = TRUE, fig.show = 'hide'}
library(ggdag)
library(ggplot2)
WageDag1 <- tidy_dagitty(dagify(
  Wage ~ Educ,
  coords = coords,
  exposure = "Educ",
  outcome = "Wage"
))
WageDag1 <-
  dag_label(WageDag1,
            labels = c("Educ" = "Exposure",
                       "Wage" = "Outcome"))
ggdag(WageDag1) + theme_dag(legend.position = 'none') + geom_dag_label_repel(aes(label = label))
```

## Omitted Variable (Confounder) Bias in a DAG

In chapter four, we learned about the problem of omitted variables bias: 

$$ln(Wage_i) = \beta_0 + \beta_1Education_i + \beta_2Experience_i + \beta_3Experience_i^2  + u_i$$

```{r, include = TRUE, fig.align = 'center'}
ggdag_collider(WageDag2) + theme_dag(legend.position = 'none') + geom_dag_label_repel(aes(label = label)) 
```

## Omitted Variable (Confounder) Bias in a DAG

```{r, echo = TRUE, fig.show = 'hide'}
WageDag2 <-
  tidy_dagitty(
    dagify(
      Wage ~ Educ + Exper,
      Educ ~ Exper,
      coords = coords,
      exposure = "Educ",
      outcome = "Wage"
    )
  )
WageDag2 <-
  dag_label(WageDag2,
            labels = c(
              "Educ" = "Exposure",
              "Wage" = "Outcome",
              "Exper" = ifelse(
                is_confounder(WageDag2, "Exper", "Educ", "Wage"),
                "Confounder",
                ""
              )
            ))
ggdag_collider(WageDag2) + theme_dag(legend.position = 'none') + geom_dag_label_repel(aes(label = label)) 
```

## Endogeneity (Collider) Bias in a DAG

Suppose we have a wage model that looks like: 

$$ln(Wage_i) = \beta_0 + \beta_1Education_i + \beta_2Experience_i + \beta_3Experience_i^2  + u_i$$
$$Education_i = \gamma_0 + \gamma_1Ability_i + v_i$$

Also suppose that ability is unobservable, but we have a variable that causes *education* but *not wages* to vary. 

```{r, include = TRUE, fig.align = 'center'}
ggdag_collider(WageDag3) + theme_dag(legend.position = 'none') + geom_dag_label_repel(aes(label = label))
```

## Endogeneity (Collider) Bias in a DAG

Now, suppose that an unobserved "latent" variable, "ability" affects both wages and education:

```{r, echo = TRUE, fig.show = 'hide'}
WageDag3 <-
  tidy_dagitty(
    dagify(
      Wage ~ Educ + Exper + Ability,
      Exper ~~ Educ,
      Educ ~ Ability + Z
      coords = coords,
      exposure = "Educ",
      outcome = "Wage",
      latent = "Ability"
    )
  )
WageDag3 <-
  dag_label(WageDag3,
            labels = c(
              "Educ" = "Exposure",
              "Wage" = "Outcome",
              "Exper" = ifelse(
                is_confounder(WageDag3, "Exper", "Educ", "Wage"),
                "Confounder", ""),
              "Ability" = "Latent", 
              "Z" = "Instrument"
            ))
WageDag3 <- control_for(WageDag3, "Ability")
ggdag_collider(WageDag3) + theme_dag(legend.position = 'none') + geom_dag_label_repel(aes(label = label))
```

## OLS and Endogeneity

Consider the following model: 

$$ln(Wage_i) = \beta_0 + \beta_1Education_i + \beta_2Experience_i + \beta_3Experience_i^2 + \beta_4Black_i$$ 
$$+ \beta_5SMSA_i + \beta_6South_i + \beta_7SMSA_i^{1966} + \sum_{r=2}^R{\delta_rRegion_{r,i}^{1966}} + u_i$$

Store the OLS estimates for this specification as $card.lm1$. 

Estimate a linear model for education as a function of whether the individual lived near a four-year college in 1966 ($nearc4$) *and* all of the other explanatory variables from the wage equation above. 

Summarize the results using $stargazer$. 

```{r card, exercise = TRUE}

```

```{r card-hint}
card.lm1 <- lm(formula, data)
card.firststage <- lm(formula, data)
stargazer(list, type)
```

```{r card-solution}
card.lm1 <- lm(log(wage) ~ educ + exper + I(exper^2) + black + smsa + south + smsa66 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669, data=card)
card.xhat <- lm(educ ~ nearc4 + exper + I(exper^2) + black + smsa + south + smsa66 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669, data = card)
stargazer(card.lm1, card.firststage, type = 'text')
```

```{r card-check}
grade_code()
```

## Implementing IV through 2SLS

Create a new named variable in $card$ called "educhat" equal to the $fitted.values$ component of the education model ($card.fiststage$) we estimated previously. 

Substitute these fitted values for the raw education variable in the wage equation. 

Use the ivreg function to calculate the "two stage least squares" (2SLS) instrumental-variables estimator for the wage equation, using proximity to a four-year college as the excluded instrument for eucation. 

Summarize the results of the OLS estimator and both of the TSLS  using $stargazer$. 

```{r cardiv, exercise = TRUE}

```

```{r cardiv-hint}
card.lm1 <- lm(formula, data)
card.firststage <- lm(formula, data)
stargazer(list, type)
```

```{r cardiv-solution}
card$educhat <- card.xhat$fitted.values
card.lm2 <- lm(log(wage) ~ educhat + exper + I(exper^2) + black + smsa + south + smsa66 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669, data=card) 
card.iv <-ivreg(log(wage) ~ educ + exper + I(exper^2) + black + smsa + south + smsa66 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 | nearc4 + exper + I(exper^2) + black + smsa + south + smsa66 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669, data = card)
stargazer(card.xhat, card.lm1, card.lm2, card.iv, type = 'text')
```

```{r cardiv-check}
grade_code()
```

Note that the IV "by hand" gives identical point estimates to $ivreg$; the standard errors differ because we have not correctly accounted for the instruments in $V(\beta) = (\hat{X}'\hat{X})^{-1}\hat\sigma_u^2$.
