---
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
title: "Appendix B"
subtitle: |
  | Fundamentals of Probability
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
description: "This tutorial practices basic concepts about relationships between two variables."
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis_setup()
tutorial_options(exercise.reveal_solution = FALSE)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Appendix B #10

Suppose that at a large university, college GPA and SAT score are related by the conditional expectation $E(GPA|SAT) = 0.70 + 0.002 SAT$
  - Find expected $GPA$ when $SAT = 800$. Find $E(GPA|SAT = 1400)$. 
  - If average $SAT$ at the university is 1,100, what is the expected average GPA? 

```{r gpaSAT, exercise=TRUE}

```

```{r gpaSAT-solution}
  0.7 + 0.002*800
  0.7 + 0.002*1100
```

```{r gpaSAT-check}
grade_this({if (identical(.result, .solution)) {
  pass(random_praise())}
  fail(random_encouragement())})
```

```{r singleObservation, echo=FALSE}
question("If an individual student's $SAT$ is 1,100, does this mean they will have the same GPA as the previous question?",
    answer("Yes"),
    answer("No", correct = TRUE),
    allow_retry = TRUE
  )
```

## Correlogram Practice

Make a correlogram of the affairs dataset that includes the following variables: 

  - Age (age)
  - Years married (yrsmarr)
  - Years of education (educ)
  - Number of affairs (naffairs)

Give your corrgram the following panel attributes: 
  
  - Scatter plots in the lower panel;
  - Correlation coefficients in the upper panel;
  - Estimated empirical uni-variate densities in the diagonal panel; 

```{r corrgramPractice, exercise=TRUE}

```

```{r corrgramPractice-hint-1}
# Use square brackets to include specific columns:  df[rows, columns]
  # by row/column number: df[c(1, 2, 3,...), c(1, 2, 3,...)]  
  # by row/column name: df(c('rowname1', 'rowname2',...), c('colname1', colname2',...))
  # leaving the row blank includes all rows: df[ , c(1,2)] for the entire first 2 rows
```

```{r corrgramPractice-hint-2}
# Use the upper.panel, lower.panel, and diag.panel to set the types of information in each part. 
  # `upper.panel = panel.cor` fpr correlation coefficients in the triangle above the diagonal. 
  # `lower.panel =`panel.pts` for scatter plots in the triangle below the diagonal. 
  # `diag.panel = 'panel.density` for a density plot (like a smoothed histogram) along the diagonal.
```

```{r corrgramPractice-hint-3}
# I used a black, monochrome (i.e. boring) color palette `col.regions = colorRampPalette(c('black'))
# Look up color options and experiment.
```

```{r corrgramPractice-solution}
library(corrgram)
affairs <- wooldridge::affairs
corrgram(affairs[, c('age', 'yrsmarr', 'educ', 'naffairs')], 
         upper.panel = panel.cor, 
         lower.panel = panel.pts, 
         diag.panel = panel.density, 
         col.regions = colorRampPalette(c('black'))) 
```

```{r corrgramPractice-check}
grade_this({
  if (!inherits(.result, c("matrix", "array"))) {
    fail("Your class of your answer should be c('matrix', 'array').")}
  if (length(.result) != 16) {
    fail("Your table should have a length of 16. Make sure you plot the variables specified.")}
  if (nrow(.result) != 4) {
    fail("Your table should have 4 rows. Make sure you plot the variables specified.")}
  pass()
})
```
