---
title: "Chapter 2"
subtitle: "The Simple Regression Model: Basic Properties"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggcheck)
gradethis_setup()
tutorial_options(exercise.reveal_solution = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

## CEO Salaries and Return on Equity

### Estimation

Regress CEO salary on:  
1. ROE (without a constant - $salary.lm0$); 
2. ROE (with a constant - name this $salary.lm1$); 
3. Only a constant($salary.lm2$). Plot all three regression lines with a scatterplot of salary on ROE. 
4. Print the `summary()` of all three regressions combined as a `list()`. 

```{r lmPractice, exercise=TRUE, exercise.reveal_solution = FALSE}

```

```{r lmPractice-hint-1}
# Don't forget to load your data!
```

```{r lmPractice-hint-2}
# The `formula` argument in `lm()` takes the form `y ~ x`
# The tilde (~) takes the place of an equals sign, separating the left from the right hand sides of the formula. 
```

```{r lmPractice-hint-3}
# Don't forget to specify your data frame!
```

```{r lmPractice-solution}
ceosal1 <- wooldridge::ceosal1
salary.lm1 <- lm(salary ~ roe, data = ceosal1)
salary.lm2 <- lm(salary ~ roe - 1, data = ceosal1)
salary.lm3 <- lm(salary ~ 1, data = ceosal1)
list(summary(salary.lm1),
     summary(salary.lm2),
     summary(salary.lm3))
```

```{r lmPractice-check}
grade_this({
  if (length(.result) == length(.solution) &
      (identical(.solution[[1]]$call, .result[[1]]$call) | 
       identical(.solution[[1]]$call, .result[[2]]$call) | 
       identical(.solution[[1]]$call, .result[[3]]$call) 
       )
      ) 
    {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Plotting the Regression Line

Plot the data for the previous exercise along with all three of the estimated regression lines.

```{r plotPractice, exercise=TRUE, exercise.reveal_solution = FALSE}

```

```{r plotPractice-setup}
ceosal1 <- wooldridge::ceosal1
salary.lm1 <- lm(salary ~ roe, data = ceosal1)
salary.lm2 <- lm(salary ~ roe - 1, data = ceosal1)
salary.lm3 <- lm(salary ~ 1, data = ceosal1)
```

```{r plotPractice-solution}
library(ggplot2)
ggplot(ceosal1) + 
  geom_point(aes(roe, salary)) + 
  geom_smooth(aes(roe, salary), method = 'lm') + 
  geom_smooth(aes(roe, salary), method = 'lm', formula = y ~ x - 1, color = 'red') + 
  geom_smooth(aes(roe, salary), method = 'lm', formula = y ~ 1, color = 'darkgreen') + 
  labs(title = "Figure 1", subtitle = "Salary and ROE", x = "ROE", y = "Salary") +
  ylim(c(0,5000))
```

```{r plotPractice-check}
grade_this({
  if (uses_geoms(.result, "point", exact = FALSE) & 
      uses_geoms(.result, "smooth", exact = FALSE) & 
      uses_mappings(.result, aes(x = roe, y = salary), exact = FALSE) & 
      .result$layers[[2]]$constructor$method == 'lm') {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```
