---
title: "Chapter 8-1: Testing for Heteroskedasticity"
author: Jim Bang
description: "Consequences of heterodkedasticity and testing for heteroskedasticity"
output: 
  learnr::tutorial: 
    progressive: true
runtime: shiny_prerendered
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

<script language="JavaScript">
$(function() {
   var editor;
   $('.ace_editor').each(function( index ) {
     editor = ace.edit(this);
     editor.setFontSize("16px");
   });
})
</script>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(lmtest)
library(sandwich)
library(car)
library(stargazer)
wage1 <- wooldridge::wage1
wage.lm14 <- lm( log(wage) ~ I(married * (1 - female)) + I(married * female) + I((1 - married) * female) + educ + exper + I(exper ^ 2) + tenure + I(tenure ^ 2),data = wage1)
wage.lm14r <- coeftest(wage.lm14, vcov = vcovHC(wage.lm14, type = "HC1"))
gpa3 <- wooldridge::gpa3
gpa.lm <- lm(cumgpa~sat+hsperc+tothrs+female+black+white, data=gpa3, subset=(spring==1))
coeftest(gpa.lm)
coeftest(gpa.lm, vcov=hccm)
hprice1 <- wooldridge::hprice1
hprice.lm1 <- lm(price ~ lotsize + sqrft + bdrms, data = hprice1)
ureg1 <- lm(resid(hprice.lm1) ^ 2 ~ lotsize + sqrft + bdrms, data = hprice1)
bpF1 <- summary(ureg1)$fstatistic
bpF1p <- pf(bpF1[1], bpF1[2], bpF1[3], lower.tail = FALSE)
bp1 <- bptest(hprice.lm1)
hprice.lm2 <- lm(log(price) ~ lotsize + sqrft + bdrms, data = hprice1)
ureg2 <- lm(resid(hprice.lm2) ^ 2 ~ lotsize + sqrft + bdrms, data = hprice1)
bpF2 <- summary(ureg2)$fstatistic
bpF2p <- pf(bpF2[1], bpF2[2], bpF2[3], lower.tail = FALSE)
bp2 <- bptest(hprice.lm2)
tutorial_options(exercise.timelimit = 60, exercise.reveal_solution = TRUE)
knitr::opts_chunk$set(echo = FALSE)
```

## Heteroskedasticity

Heteroskedasticity does *not* affect the unbiased-ness or consistency of $\hat\beta_{OLS}$. 

Heteroskedasticity *does* affect the unbiased-ness of the standard errors of $\hat\beta_{OLS}$, $\hat\sigma_{\hat\beta_{OLS}}$. 

Inferences and interval estimates for $\beta$ will therefore be wrong. 

### Heteroskedasticity-Robust Inference for $log(wage)$ Model

Run the following code to replicate example 8.1 from the book. 

```{r robustHC1, exercise = TRUE}
wage.lm14 <-
  lm(
    log(wage) ~ I(married * (1 - female)) + I(married * female) + 
      I((1 - married) * female) + educ + exper + I(exper ^ 2) + tenure +  
      I(tenure ^ 2),data = wage1
  )
wage.lm14r <-
  coeftest(wage.lm14, vcov = vcovHC(wage.lm14, type = "HC1"))
stargazer(wage.lm14, wage.lm14r, type = 'text', digits = 4)
```

```{r robustHC1-solution}
wage.lm14 <-
  lm(
    log(wage) ~ I(married * (1 - female)) + I(married * female) + 
      I((1 - married) * female) + educ + exper + I(exper ^ 2) + tenure + 
      I(tenure ^ 2), data = wage1
  )
wage.lm14r <-
  coeftest(wage.lm14, vcov = vcovHC(wage.lm14, type = "HC1"))
stargazer(wage.lm14, wage.lm14r, type = 'text', digits = 4)
```

```{r robustHC1-check}
grade_this({if (!inherits(.result, c("character"))) {
    fail("Your class of your answer should be a character list")}
  if (.result[10] != "I(married * (1 - female))        0.2127***          0.2127*** ") { 
    fail("The coefficient on married women don't look quite right.") }
  pass("Note that the coefficients are the same; only the standard errors differ.")
})
```

### Heteroskedasticity-Robust F-Test for College GPAs

Run the following code to replicate example 8.2 from the book. 

```{r robustF, exercise = TRUE}
collGPA.lm <- lm(
  cumgpa ~ sat + hsperc + tothrs + female + black + white, data = gpa3,
  subset = (spring == 1)
)
collGPA.lmr <-
  coeftest(collGPA.lm, vcov = vcovHC(collGPA.lm, type = "HC1"))
fStat <- linearHypothesis(collGPA.lm, c('black', 'white'))
fStat.r <-
  linearHypothesis(collGPA.lm, c('black', 'white'), 
                   vcov = hccm(collGPA.lm, type = "hc0"))
stargazer(
  collGPA.lm,
  collGPA.lmr,
  type = 'text',
  digits = 4,
  add.lines = list(
    c('F-Test (Race)', round(fStat$F[2], 4), round(fStat.r$F[2], 4)),
    c('Pr(>F)', round(fStat$`Pr(>F)`[2], 4), round(fStat.r$`Pr(>F)`[2], 4))
  )
)
```

```{r robustF-solution}
collGPA.lm <- lm(
  cumgpa ~ sat + hsperc + tothrs + female + black + white, data = gpa3,
  subset = (spring == 1)
)
collGPA.lmr <-
  coeftest(collGPA.lm, vcov = vcovHC(collGPA.lm, type = "HC1"))
fStat <- linearHypothesis(collGPA.lm, c('black', 'white'))
fStat.r <-
  linearHypothesis(collGPA.lm, c('black', 'white'), 
                   vcov = hccm(collGPA.lm, type = "hc0"))
stargazer(
  collGPA.lm,
  collGPA.lmr,
  type = 'text',
  digits = 4,
  add.lines = list(
    c('F-Test (Race)', round(fStat$F[2], 4), round(fStat.r$F[2], 4)),
    c('Pr(>F)', round(fStat$`Pr(>F)`[2], 4), round(fStat.r$`Pr(>F)`[2], 4))
  )
)
```

```{r robustF-check}
grade_this({if (!inherits(.result, c("character"))) {
    fail("Your class of your answer should be a character list")}
  if (.result[32] != "F-Test (Race)                0.6796            0.7478   ") { 
    fail("The line that should contain the partial F test doesn't look quite right.") }
  pass()
})
```

## Testing for Heteroskedasticity  

Heteroskedasticity tests involve estimating the model by OLS followed by estimating the residuals as a function of the explanatory variables from that model. 

### Breusch-Pagan Test

1. Estimate the model of interest using OLS. 
2. Regress the squared residuals from (1) on the $x$ variables from your model. 
3. Calculate the $F$ statistic *or* the $LM$ statistic for this regression. <br>
$F = \frac{R_{\hat{u}^2}^2/k}{(1-R_{\hat{u}^2}^2)/(n-k-1)}$<br>
$LM = nR_{\hat{u}^2}^2$
4. Test the statistic you calculated on $F \sim F(k, n-k-1)$ or $LM \sim \chi^2(k)$

### House Prices

1. Estimate `hprice.lm1` as the model of house prices given in by <br>
$Price = \beta_0 + \beta_1LotSize + \beta_2SquareFeet + \beta_3Bedrooms + u$
2. Estimate the regression of the squared residuals on the explanatory variables (`ureg1`). 
3. Define the Breusch-Pagan $F$-Statistic (`bpF1`) as the F-Statistic of the residual regression and the corresponding p-value (`bpF1p`) as the probability in the *upper* tail greater than F. 
4. Calculate the Breusch-Pagan $LM$-Statistic using `bptest`. 
5. Repeat 1-4 using $log(Price)$ as the dependent variable. 
6. I have built a `stargazer` table for you (you're welcome!). Notice my use of the `add.lines` option to insert the specification tests for heteroskedasticity. 

```{r bpTest, exercise = TRUE}
# Levels Model

# Log Model

# Table
stargazer(hprice.lm1, hprice.lm2, type = 'text', digits = 4,
  add.lines = list(
    c('Breusch-Pagan F', round(bpF1[1], 4), round(bpF2[1], 4)),
    c('Pr(>F)', round(bpF1p, 4), round(bpF2p, 4)),
    c('Breusch-Pagan LM', round(bp1$statistic, 4), round(bp2$statistic, 4)),
    c('Pr(>Chi-Sq)', round(bp1$p.value, 4), round(bp2$p.value, 4)
    )
  )
)
```

```{r bpTest-hint1}
# Let `bpF1` be the `fstatistic` object from `summary(ureg1)`
```

```{r bpTest-hint2}
# The `fstatistic` object has three values: (1) the F-stat itself, (2) the numerator df, and (3) the denominator df. Keep all three in `bpF1`. 
```

```{r bpTest-hint3}
# Use `pf(F, df1, df2)` to calculate the p-value for F using the three values from `fstatisitc`.  
```

```{r bpTest-solution}
# Levels Model
hprice.lm1 <- lm(price ~ lotsize + sqrft + bdrms, data = hprice1)
ureg1 <- lm(resid(hprice.lm1) ^ 2 ~ lotsize + sqrft + bdrms, data = hprice1)
bpF1 <- summary(ureg1)$fstatistic
bpF1p <- pf(bpF1[1], bpF1[2], bpF1[3], lower.tail = FALSE)
bp1 <- bptest(hprice.lm1)
# Log Model
hprice.lm2 <- lm(log(price) ~ lotsize + sqrft + bdrms, data = hprice1)
ureg2 <- lm(resid(hprice.lm2) ^ 2 ~ lotsize + sqrft + bdrms, data = hprice1)
bpF2 <- summary(ureg2)$fstatistic
bpF2p <- pf(bpF2[1], bpF2[2], bpF2[3], lower.tail = FALSE)
bp2 <- bptest(hprice.lm2)
# Table
stargazer(hprice.lm1, hprice.lm2, type = 'text', digits = 4,
  add.lines = list(
    c('Breusch-Pagan F', round(bpF1[1], 4), round(bpF2[1], 4)),
    c('Pr(>F)', round(bpF1p, 4), round(bpF2p, 4)),
    c('Breusch-Pagan LM', round(bp1$statistic, 4), round(bp2$statistic, 4)),
    c('Pr(>Chi-Sq)', round(bp1$p.value, 4), round(bp2$p.value, 4)
    )
  )
)
```

```{r bpTest-check}
grade_this({if (!inherits(.result, c("character"))) {
    fail("Your class of your answer should be a character list")}
  if (.result[21] != "Breusch-Pagan F                   5.3389        1.1745    ") { 
    fail("The line that should contain the Breusch-Pagan F test doesn't look quite right.") }
  pass("Wait, what happened to the heteroskedasticity in the log-linear model?!?")
})
```

### White Test

The White test for heteroskedasticity is similar to the Breusch-Pagan test, except it includes all possible interactions among the variables, including with themselves (i.e. quadratic terms). 

1. Estimate the model.
2. Estimate $\hat{u}^2 = \delta_0 + \delta_1x_1 + \delta_2x_2 + \delta_3x_1^2 + \delta_4x_2^2 + \delta_5x_1x_2 + v$, <br> 
which is equivalent to $\hat{u}^2 = \delta_0 + \delta_1\hat{y} + \delta_2\hat{y}^2 + v$.
3. Calculate $F \sim F(2, n-3)$ or $LM \sim \chi^2(2)$ as you did in the Breusch-Pagan test. 

Note: none of the degrees of freedom in $F$ or $LM$ depend on $k$ because you always only estimate three parameters in the squared-residuals regression. 

### House Prices

Calculate the White test for the housing models you conducted Breusch-Pagan tests for in the previous exercise. You should not need to re-estimate the regression models for house prices (`hprice.lm1` and `hprice.lm2`). 

```{r white, exercise = TRUE}
# Levels Model

# Log Model

# Table
stargazer(hprice.lm1, hprice.lm2, type = 'text', digits = 4,
  add.lines = list(
    c('White F', round(whiteF1[1], 4), round(whiteF2[1], 4)),
    c('Pr(>F)', round(whiteF1p, 4), round(whiteF2p, 4)),
    c('White LM', round(white1$statistic, 4), round(white2$statistic, 4)),
    c('Pr(>Chi-Sq)', round(white1$p.value, 4), round(white2$p.value, 4))))
```

```{r white-solution}
ureg1w <-
  lm(resid(hprice.lm1) ~ fitted(hprice.lm1) + I(fitted(hprice.lm1) ^ 2))
whiteF1 <- summary(ureg1w)$fstatistic
whiteF1p <- pf(whiteF1[1], whiteF1[2], whiteF1[3], lower.tail = FALSE)
white1 <-
  bptest(hprice.lm1, ~ fitted(hprice.lm1) + I(fitted(hprice.lm1) ^ 2))
ureg2w <-
  lm(resid(hprice.lm2) ~ fitted(hprice.lm2) + I(fitted(hprice.lm2) ^ 2))
whiteF2 <- summary(ureg2w)$fstatistic
whiteF2p <- pf(whiteF2[1], whiteF2[2], whiteF2[3], lower.tail = FALSE)
white2 <-
  bptest(hprice.lm2, ~ fitted(hprice.lm2) + I(fitted(hprice.lm2) ^ 2))
stargazer(hprice.lm1, hprice.lm2, type = 'text', digits = 4,
  add.lines = list(
    c('White F', round(whiteF1[1], 4), round(whiteF2[1], 4)),
    c('Pr(>F)', round(whiteF1p, 4), round(whiteF2p, 4)),
    c('White LM', round(white1$statistic, 4), round(white2$statistic, 4)),
    c('Pr(>Chi-Sq)', round(white1$p.value, 4), round(white2$p.value, 4))))
```

```{r white-check}
grade_this({if (!inherits(.result, c("character"))) {
    fail("Your class of your answer should be a character list")}
  if (.result[21] != "Breusch-Pagan F                   5.3389        1.1745    ") { 
    fail("The line that should contain the White F test doesn't look quite right.") }
  pass("Wait, what happened to the heteroskedasticity in the log-linear model?!?")
})
```
