---
title: "Appendix B-2 Practice"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis_setup()
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```

## Appendix B #10

Suppose that at a large university, college GPA and SAT scoreare related by the conditional expectation $E(GPA|SAT) = 0.70 + 0.002 SAT$
  - Find expected $GPA$ when $SAT = 800$. Find $E(GPA|SAT = 1400)$. 
  - If average $SAT$ at the university is 1,100, what is the average GPA? 

```{r gpaSAT, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r gpaSAT-hint}

```

```{r gpaSAT-solution}

```

```{r gpaSAT-check}
grade_result() 
```

```{r singleObservation, echo=FALSE}
question("If a student's $SAT$ is 1,100, does this mean they will have the same GPA as the previous question?",
    answer("Yes"),
    answer("No", correct = TRUE),
    allow_retry = TRUE
  )
```


## Correlogram Practice

Make a correlogram of the affairs dataset that includes the following variables: 

  - Age (age)
  - Years married (yrsmarr)
  - Years of education (educ)
  - Number of affairs (naffairs)

Give your corrgram the following panel attributes: 
  
  - Scatter plots in the lower panel;
  - Correlation coefficients in the upper panel;
  - Estimated empirical uni-variate densities in the diagonal panel; 
  - Black color palette. 

```{r corrgramPractice, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r corrgramPractice-hint}

```

```{r corrgramPractice-solution}
corrgram(affairs[,which(names(affairs) %in% c('age', 'yrsmarr', 'educ', 'naffairs'))], 
         upper.panel = panel.cor, 
         lower.panel = panel.pts, 
         diag.panel = panel.density, 
         col.regions = colorRampPalette(c('black'))) 
```

```{r corrgramPractice-check}
grade_result() 
```
