---
title: "Chapter 1-1 Practice"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis_setup()
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```

## Exercise

Here's a simple exercise with an empty code chunk provided for entering the answer.

Write the R code required to perform the following operations:

- $21 + 21$
- $5(4-1)^2-3$
- $\sqrt{1764}$

```{r R_as_a_calculator, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r R_as_a_calculator-solution}
21+21
5*(4-1)^2-3
sqrt(1764)
```

```{r R_as_a_calculator-hint}
# R can be used as a calculator using the command line e.g.:
1 + 1
```

```{r R_as_a_calculator-check}
grade_result(
  pass_if(~identical(.result, 42))) 
```

## Using Objects

Do the following: 

- Assign a value of five to a numerical object x
- Display x and it's squared value
- Assign a value of 3 to the object y and a value of $y^x$ to the object z using these objects.

```{r objects, exercise=TRUE, exercise.eval=TRUE}

```

```{r objects-solution}
x <- 5
x
x^2
y <- 3
z <- y^x
```

```{r objects-code-check}
grade_code() 
```

## Vectors and Matrices

Do the following: 

- Create the vector $a = \begin{pmatrix} 1 \\ 2 \\ 3 \\ 4 \\ 5 \\ 6 \end{pmatrix}$;
- Assign the value $a + 1$ to the object b;
- Assign the value $a + b$ to the object c;
- Assign the value $b + c$ to the object d;
- Calculate $\sqrt{d}$;

```{r Vectors, exercise=TRUE, exercise.eval=TRUE}

```

```{r Vectors-solution}
a <- c(1,2,3,4,5,6)
b <- a+1
c <- a+b
d <- b+c
sqrt(d)
A <- matrix(a, ncol = 2)

```

```{r Vectors-code-check}
grade_code() 
```

- Create a matrix, A that splits the values of a across two columns;
- Create a matrix, B that duplicates the columns of A twice. 
- Extract the third column of B as a matrix named "b."

```{r Matrices, exercise=TRUE, exercise.eval=TRUE}

```

```{r Matrices-solution}
A <- matrix(a, ncol = 2)
B <- cbind(A, A)
b <- B[, 3]
```

```{r Matrices-code-check}
grade_code() 
```

## Loading and importing data frames

Load the RData file, "BeerData" into your workspace, then import the CSV file by the same name as an object named "BeerDataCSV"

```{r Import, exercise=TRUE, exercise.eval=TRUE}

```

```{r Import-solution}
load("BeerData.RData")
BeerDataCSV <- read.csv("BeerData.csv")
```

```{r Import-code-check}
grade_code() 
```

## Saving your work

When you are finished it is *crucial* that you save your work. Save the object, BeerDataCSV as its own RData file named "BeerDataCSV.RData," then save your entire workspace as "BeerData.RData."

```{r Save, exercise=TRUE, exercise.eval=TRUE}

```

```{r Save-solution}
save(BeerDataCSV, file = "BeerDataCSV.RData")
save.image("BeerData.RData")
```

```{r Save-code-check}
grade_code() 
```

## Examining a Dataset
 
Display the first *five* rows of the mtcars dataset. Then, View the complete data frame in a new window.

```{r View, exercise=TRUE, exercise.eval=TRUE}

```

```{r View-solution}
head(mtcars, n = 5)
View(mtcars)
```

```{r View-code-check}
grade_code() 
```

## Writing your own function 

Now write a function that adds any two numbers, x and y:

```{r add_function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

```{r add_function-hint}
# A function needs arguments (variables) as inputs.
function(...)
```


```{r add_function-solution}
add <- function(x, y) {
  x + y
}
```

```{r add_function-code-check}
grade_code() 
```

