---
title: "Chapter 1-1 Practice"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis_setup()
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```

## Exercise

Here's a simple exercise with an empty code chunk provided for entering the answer.

Write the R code required to perform the following operations:

- $21 + 21$
- $5(4-1)^2-3$
- $\sqrt{1764}$

```{r R_as_a_calculator, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r R_as_a_calculator-solution}
21+21
5*(4-1)^2-3
sqrt(1764)
```

```{r R_as_a_calculator-hint}
# R can be used as a calculator using the command line e.g.:
1 + 1
```

```{r R_as_a_calculator-check}
grade_result(
  pass_if(~identical(.result, 42))) 
```

### Using Objects

R is an object-oriented language. The "assignment" operator, "<-" (keyboard shortcut: "Alt" + "-") computes the operation to the right and gives it name in your *environment* (see top-right pane in RStudio). Name the calculations from the previous exercise as indicated below.

- $21 + 21$: answer
- $5(4-1)^2-3$: life
- $\sqrt{1764}$: universe
- $ln(e^{42})$: everything

Display the objects in your environment using the `ls()` function. 

```{r objects, exercise=TRUE}

```

```{r objects-solution}
answer <- 21+21
life <- 5*(4-1)^2-3
universe <- sqrt(1764)
everything <- log(exp(42))
ls()
```

```{r objects-check}
grade_code() 
```

Note that in some cases you can use the "equal to" operator, "=", but only if the thing you're operating and the name of the object you're assigning are unambiguous. The direction of the "<-" operator points *from* the operation you mean for R to calculate and *to* the name you want to give to it. In fact, you may reverse the direction to assign from left to right, as in `21 + 21 -> answer`. 

### Defining and Evaluating Functions

If you can define an object you can write (and evaluate) your own function! A function can be a simple mathematical function, or it can be a long script that outputs several different values. Define the function, `quadratic`, as a function of *variable* `x` and *parameters* `a`, `b`, and `c` as 
$$f(x) = ax^2 + bx + c.$$ 
Then, evaluate $f(x) = 8x^2 - 6x - 12$ at $x = 3$

```{r functions, exercise=TRUE}

```

```{r functions-solution}
quadratic <- function(x, a, b, c) {
  a*x^2 + b*x + c
}
quadratic(3, 8, -6, -12)
```

```{r functions-check}
grade_code() 
```

For the purposes of getting computational solutions, there is not really a difference between a parameter and a variable. Economically, we consider parameters (a, b, c, and d) to be *exogenous* - determined by nature (like preference parameters) or markets (like prices or income) - whereas we consider variables (x) to be *endogenous* - chosen by households (like quantities consumed) or firms (like amounts of a resource hired). 

## Vectors, Matrices, and Data Frames

### Vectors

Do the following: 

- Create the vector $a = \begin{pmatrix} 1 \\ 2 \\ 3 \\ 4 \\ 5 \\ 6 \end{pmatrix}$;
- Assign the value $a + 1$ to the object b;
- Assign the value $a + b$ to the object c;
- Assign the value $b + c$ to the object d;
- Calculate $\sqrt{d}$;

```{r Vectors, exercise=TRUE, exercise.eval=TRUE}

```

```{r Vectors-solution}
a <- c(1,2,3,4,5,6)
b <- a+1
c <- a+b
d <- b+c
sqrt(d)
```

```{r Vectors-code-check}
grade_code() 
```

### Matrices

Do the following: 

- Create a matrix, A that splits the values of a across two columns;
- Create a matrix, B that duplicates the columns of A twice. 
- Extract the third column of B as a matrix named "b."

```{r Matrices, exercise=TRUE, exercise.eval=TRUE}

```

```{r Matrices-solution}
A <- matrix(a, ncol = 2)
B <- cbind(A, A)
b <- B[, 3]
```

```{r Matrices-code-check}
grade_code() 
```

### Data Frames

1. *Load* "BeerData.RData" from "https://github.com/bangecon/metricsToTheFace/tree/master/inst/tutorials/Chapter01-1-Practice/BeerData.RData" into your workspace. 
2. *Read* "BeerData.csv" from "https://github.com/bangecon/metricsToTheFace/tree/master/inst/tutorials/Chapter01-1-Practice/BeerData.csv" as an object named "BeerDataCSV".

```{r Import, exercise=TRUE, exercise.eval=TRUE}

```

```{r Import-solution}
load("https://github.com/bangecon/metricsToTheFace/tree/master/inst/tutorials/Chapter01-1-Practice/BeerData.RData")
BeerDataCSV <- read.csv("https://github.com/bangecon/metricsToTheFace/tree/master/inst/tutorials/Chapter01-1-Practice/BeerData.csv")
```

```{r Import-code-check}
grade_code() 
```

### Examining a Dataset
 
Display the first *five* rows of the mtcars dataset. Then, View the complete data frame in a new window.

```{r View, exercise=TRUE, exercise.eval=TRUE}

```

```{r View-solution}
head(mtcars, n = 5)
View(mtcars)
```

```{r View-code-check}
grade_code() 
```

