---
title: "Appendix C \n\n Fundamentals of Mathematical Statistics"
author: "James Bang"
description: >
  This tutorial introduces general concepts of estimation and testing.
editor_options: 
  markdown: 
    wrap: 72
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

<script language="JavaScript">
$(function() {
   var editor;
   $('.ace_editor').each(function( index ) {
     editor = ace.edit(this);
     editor.setFontSize("16px");
   });
})
</script>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
audit <- wooldridge::audit
avgy<- mean(audit$y)
n   <- length(audit$y)
sdy <- sd(audit$y)
se  <- sdy/sqrt(n)
c05   <- abs(qt(.025, n-1))
c01   <- abs(qt(.005, n-1))
gradethis::gradethis_setup(exercise.reveal_solution = TRUE)
knitr::opts_chunk$set(echo = FALSE)
```

## Approaches to Parameter Estimation

### Ordinary Least Squares (OLS)

$$\min_{\hat{\beta}}{\sum_{i=1}^{n}(Y_i – \beta_0 – \beta_1X_i)^2}$$

### Least/Minimum Absolute Deviations (LAD/MAD)

$$\min_{\hat{\beta}}{\sum_{i=1}^{n}|Y_i – \beta_0 – \beta_1X_i|}$$

### Maximum likelihood (MLE)

$$\max_{\hat{\beta}}{\prod_{i=1}^{n}f(Y_i – \beta_0 – \beta_1X_i)}$$

where $f(\cdot)$ is the probability distribution of the errors (e.g. Normal, logistic, Poisson)

### Method of Moments 

$$E(Xu) = 0$$
$$ \sum_{i=1}^n{X_i(Y_i-\beta_0-\beta_1X_i)} = 0 $$ 

If the assumptions of the Classical Regression Model hold, all four of these methods are equivalent.

## Interval Estimation

The 95% confidence interval for the sample mean, $\bar{x}$ solves
$$ P(\bar{x}-t_{0.025}^c \cdot s_{\bar{x}} \le \mu \le \bar{x}+t_{0.025}^c \cdot s_{\bar{x}}) = 0.95 $$
Since the sampling distribution of $\bar{x}$ is normal and the sampling
distribution of $s_{\bar{x}}^2$ is $\chi^2$, this involves inverting the
t-distribution (normal divided by the square root of $\chi^2$).

### Example

Using the $audit$ data from the $wooldridge$, calculate the 95% confidence interval *by hand* by calculating the following (and naming them): 

1. the mean of `y` (call it `avgy`);
2. the number of observations in (`length()` of) `y` (`n`);
3. the standard deviation of `y` (`sdy`)
4. the standard error of $\bar{y}$ (`sdy`)
5. the two-sided critical value for $\alpha = 0.05$ (`c05`) 
6. the 95% confidence-interval bounds as a vector (lower bound, upper bound)

```{r intervalEst, exercise = TRUE}
# Ingredients to CI formula: avgy, n, sdy, se, c05


# 95% Confidence interval:

```

```{r intervalEst-solution}
avgy<- mean(audit$y)
n   <- length(audit$y)
sdy <- sd(audit$y)
se  <- sdy/sqrt(n)
c05   <- abs(qt(.025))
avgy + c05 * c(-se,+se)
```

```{r intervalEst-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Hypothesis Testing

1. State the Null & Alternative Hypotheses
2. Determine the Significance Level ($\alpha$)
3. Calculate the parameters and the test statistic
4. Calculate the critical value or p-value
5. Make a Rejection Decision

### Example

Calculate the test statistic, absolute critical value for $\alpha = 0.01$, and p-value to test the null hypothesis of $H_0: \mu_y = 0$ against the two-sided alternative of $H_1: \mu_y \ne 0$. Call these t, c01, and p, and print them as a vector.

```{r tTest, exercise = TRUE}
# t statistic for H0: mu=0

# critical value for alpha = 0.05

# p value

```

```{r tTest-solution}
t <- avgy/se
c01 <- abs(qt(0.005, n-1))
p <- pt(t, n-1)
c(t, c01, p)
```

```{r tTest-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Shortcut using $t.test()$

Replicate the confidence interval and t-test from the previous examples using the `t.test()` command. Don't forget that the second example used a different $\alpha$ than the first one. 

```{r t-test, exercise = TRUE}


```

```{r t-test-solution}
t.test(audit$y)
t.test(audit$y, conf.level = 0.99)
```

```{r t-test-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```
