---
title: "Chapter 8" 
subtitle: "Testing for Heteroskedasticity"
description: "This tutorial explains the consequences of heterodkedasticity and how to test for it."
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = FALSE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```

## Net Total Financial Wealth

Use the `k401ksubs` dataset to estimate the effect of income (`inc`) on total net financial assets (`nettfa`) for *single-member households only* (`fsize == 1`) with simple regression analysis and multiple regression analysis, controlling for gender (`male`), 401k plan eligibility (`e401k`), and the square of the difference between a person's current age and 25 years old $(`age` - 25)^2$. 

Present your results as a formatted text table with `stargazer()`.

```{r nettfa, exercise = TRUE}

```

```{r nettfa-setup}

```

```{r nettfa-hint}
# Remember to use the as-is function, `I()`, for squaring `age` - 25. 
# Select single households by setting the option `subset = (variable.name == value)` option. 
```

```{r nettfa-solution}
library(stargazer)
k401ksubs <- wooldridge::k401ksubs
nettfa.lm1 <- lm(nettfa ~ inc, data = k401ksubs, subset = (fsize == 1))
nettfa.lm2 <- lm(nettfa ~ inc + male + e401k + I((age - 25)^2), data = k401ksubs, subset = (fsize == 1))
stargazer::stargazer(nettfa.lm1, nettfa.lm2, type = 'text')
```

```{r nettfa-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Breusch-Pagan Test

For the net financial assets regression *with controls*, calculate the Breusch-Pagan F-statistic (by hand), and LM statistic (howeverso you like) with the corresponding p-values. (The `lm()` estimates for the specification that includes controls are stored as `nettfa.lm2`.)  

Print your response as a list consisting of (1) a vector containing the Breusch-Pagan F-stat (including degrees of freedom) and p-value; and (2) the Breusch-Pagan LM test result from `bptest()`.

```{r bp, exercise = TRUE}

```

```{r bp-setup}
library(stargazer)
k401ksubs <- wooldridge::k401ksubs
nettfa.lm1 <- lm(nettfa ~ inc, data = k401ksubs, subset = (fsize == 1))
nettfa.lm2 <- lm(nettfa ~ inc + male + e401k + I((age - 25)^2), data = k401ksubs, subset = (fsize == 1))
```

```{r bp-hint}

```

```{r bp-solution}
library(lmtest)
ureg1 <- lm(resid(nettfa.lm2) ^ 2 ~ inc + male + e401k + I((age - 25)^2), data = subset(k401ksubs, fsize == 1))
bpF1 <- summary(ureg1)$fstatistic
bpF1p <- pf(bpF1[1], bpF1[2], bpF1[3], lower.tail = FALSE)
bp1 <- bptest(nettfa.lm2)
list(c(bpF1, bpF1p), bp1)
```

```{r bp-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## White Test

For the net financial assets regression *with controls*, calculate the White F-statistic (by hand), and LM statistic (howeverso you like) with the corresponding p-values. (The `lm()` estimates for the specification that includes controls are stored as `nettfa.lm2`.)  

Present your result as a list just like the previous example. 

```{r white, exercise = TRUE}

```

```{r white-setup}
library(stargazer)
k401ksubs <- wooldridge::k401ksubs
nettfa.lm1 <- lm(nettfa ~ inc, data = k401ksubs, subset = (fsize == 1))
nettfa.lm2 <- lm(nettfa ~ inc + male + e401k + I((age - 25)^2), data = k401ksubs, subset = (fsize == 1))
library(lmtest)
ureg1 <- lm(resid(nettfa.lm2) ^ 2 ~ inc + male + e401k + I((age - 25)^2), data = subset(k401ksubs, fsize == 1))
bpF1 <- summary(ureg1)$fstatistic
bpF1p <- pf(bpF1[1], bpF1[2], bpF1[3], lower.tail = FALSE)
bp1 <- bptest(nettfa.lm2)
list(c(bpF1, bpF1p), bp1)
```

```{r white-hint}

```

```{r white-solution}
ureg1w  <- lm(resid(nettfa.lm2) ^ 2 ~ fitted(nettfa.lm2) + I(fitted(nettfa.lm2)^2))
whiteF  <- summary(ureg1w)$fstatistic
whiteFp <- pf(whiteF[1], whiteF[2], whiteF[3], lower.tail = FALSE)
white   <- bptest(nettfa.lm2, ~ fitted(nettfa.lm2) + I(fitted(nettfa.lm2)^2), 
                 data = subset(k401ksubs, fsize == 1))
list(c(whiteF, whiteFp), white)
```

```{r white-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

Note: Since *net* total financial assets can (and do!) take negative values, we cannot simply fix the problem with logs (to the extent that taking logs really "fixes" anything.
