---
title: "Chapter 6-1 Practice"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(stargazer)
gpa2 <- wooldridge::gpa2
colgpa.lm1 <- lm(colgpa ~ athlete + female + I(1-white) + hsize + I(hsize^2) + hsperc + sat, data = gpa2)
colgpa.lm2 <- lm(colgpa ~ athlete + female + I(athlete*female) + I(1-white) + hsize + I(hsize^2) + hsperc + sat, data = gpa2)
apple <- wooldridge::apple
apple.lm1 <- lm(ecobuy ~ ecoprc + regprc + faminc + hhsize + educ + age, data = apple)
gradethis_setup()
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```

## Exercise

Estimate the following model using the gpa2 dataset: 

$$College\text{ }GPA = \beta_0 + \beta_1Athlete + \beta_2Female + \beta_3HS\text{ }Size + \beta_4HS\text{ }Size^2 + \beta_5HS\text{ }Percentile + \beta_6SAT + \beta_7Nonwhite \text{ : (gpa.lm1)} $$ 

Estimate the same model with an interaction between athlete and gender. Present your results side-by-side using $stargazer()$. 

Finally, estimate the model with an interaction between athlete and high school percentile. 

```{r athleteGPAs, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r athleteGPAs-solution}
colgpa.lm1 <- lm(colgpa ~ athlete + female + I(1-white) + hsize + I(hsize^2) + hsperc + sat, data = gpa2)
colgpa.lm2 <- lm(colgpa ~ athlete + female + I(athlete*female) + I(1-white) + hsize + I(hsize^2) + hsperc + sat, data = gpa2)
colgpa.lm3 <- lm(colgpa ~ athlete + female + I(athlete*scale(hsperc, scale = FALSE)) + I(1-white) + hsize + I(hsize^2) + hsperc + sat, data = gpa2)
stargazer(colgpa.lm1, colgpa.lm2, colgpa.lm3, type = 'text')
```

```{r athleteGPAs-hint}
# Don't forget to use the I() operator in the terms that require math operations.
# Don't forget to scale continuous interactions. 
gpa.lm1 <- lm(..., ...)
gpa.lm2 <- lm(..., ...)
gpa.lm3 <- lm(..., ...)
stargazer(gpa.lm1, gpa.lm2, gpa.lm3, ...)
```

```{r athleteGPAs-check}
grade_code()
```

## Exercise

Estimate the following model using the $apple$ dataset: 

$$ Buy\text{ }Eco\text{-}Freindly = \beta_0 + \beta_1Price_{Eco} + \beta_2Price_{Reg} + \beta_3Income + \beta_4HH\text{ }Size + \beta_5Education + \beta_6Age + u $$ 

Summarize this estimation using the $summary()$ function. 

```{r apple, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r apple-solution}
apple.lm1 <- lm((ecolbs > 0) ~ ecoprc + regprc + faminc + hhsize + educ + age, data = apple)
summary(apple.lm1)
```

```{r apple-hint}
# Note that you can create a dummy variable for values greater, less than, or equal to a threshold value using the <, >, and = operators within parentheses. You do not need the I() function, just parentheses. 
apple.lm1 <- lm(..., ...)
summary(...)
```

```{r apple-check}
grade_code()
```

Interpret the coefficient on the price of eco-friendly apples. 
