---
output: learnr::tutorial
runtime: shiny_prerendered
title: "Chapter 13 \n\n Pooling Cross Sections across Time: \n\n Simple Panel Data Methods"
author: "James Bang"
description: >
  This tutorial provides practice with pooled cross sections and difference-in-difference methods.
---

```{r setup, include=FALSE}
library(learnr)
library(wooldridge)
library(stargazer)
fertil1 <- wooldridge::fertil1
fertil1$region <- factor(fertil1$east + 2*fertil1$northcen + 3*fertil1$west, 
                         labels = c("South Central", "East", "North Central", "West"))
fertil1$geog <- factor(fertil1$farm + 2*fertil1$othrural + 3*fertil1$town + 4*fertil1$smcity, 
                       labels = c("City", "Farm", "Other Rural", "Town", "Small City"))
fertil.lm1 <- lm(kids ~ educ + age + I(age^2) + black + region + geog, data = fertil1)
fertil.lm2 <- lm(kids ~ educ + age + I(age^2) + black + region + geog + factor(year), data = fertil1)
injury <- wooldridge::injury
injury$injtype <- factor(injury$injtype, 
                         labels = c("Head", "Neck", "Upper Extremeties","Trunk", "Lower Back", 
                                    "Lower Extremities", "Occupational Disease", "Other"))
injury.lm1 <- lm(log(durat) ~ afchnge*highearn, data = injury)
injury.lm2 <- lm(log(durat) ~ afchnge*highearn + injtype + male, data = injury)
knitr::opts_chunk$set(echo = FALSE)
```

## Pooled Cross Sections: Fertility over Time

Estimate the determinants of women's fertility using the $fertil1$ data, controlling for age, age-squared, race (black vs. non-black), region, and geographic type (city, farm, rural, town, small city). Name this $fertil.lm1$.

Next, estimate the "non parametrice time trend" (year dummies) for fertililty over time. Name This $fertil.lm2$.

Summarize both regressions using a "stargazer" text-table.

```{r fertility, exercise = TRUE}

```

```{r fertility-hint}
fertil.lm1 <- lm(formula, data)
fertil.lm2 <- lm(formula, data)
stargazer(lm-object, ..., type)
```

```{r fertility-solution, exercise.reveal_solution = FALSE}
fertil.lm1 <- lm(kids ~ educ + age + I(age^2) + black + region + geog, data = fertil1)
fertil.lm2 <- lm(kids ~ educ + age + I(age^2) + black + region + geog + factor(year), data = fertil1)
stargazer(fertil.lm1, fertil.lm2, type = "text")
```

```{r fertility-check}
grade_code()
```

## Difference-in-Difference: Worker Compensation and Weeks Out-of-Work

Estimate the effect on worker's compensation duration from a change in worker's compensation that raised the earnings cap on benefits. The simple DiD model for this would look like:

$$Duration_i = \beta_0 + \delta_0After Change + beta_1HighEarner_i + \delta_1AfterChanges*HighEarner_i +  u_{it}$$ Name this $workcomp.lm1$.

Next, estimate the same model while also controlling for gender ($male$) and type of injury ($injtype$). Name this specification $workcomp.lm2$.

Summarize both regressions using a "stargazer" text-table.

```{r workcomp, exercise = TRUE}

```

```{r workcomp-hint}
workcomp.lm1 <- lm(formula, data)
workcomp.lm1 <- lm(formula, data)
stargazer(lm-object, ..., type)
```

```{r workcomp-solution}
workcomp.lm1 <- lm(log(durat) ~ afchnge*highearn, data = injury)
workcomp.lm1 <- lm(log(durat) ~ afchnge*highearn + male + injtype, data = injury)
stargazer(fertil.lm1, fertil.lm2, type = "text")
```

```{r workcomp-check}
grade_code()
```
