---
title: "Chapter 6-2 Practice"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(stargazer)
hprice1 <- wooldridge::hprice1
hprice.lm1 <- lm(price ~ lotsize + sqrft + bdrms, data = hprice1)
hprice.lm2 <- lm(price ~ lotsize + sqrft + bdrms + I(sqrft*bdrms), data = hprice1)
hprice.lm3 <- lm(price ~ lotsize + sqrft + bdrms + I((sqrft-mean(sqrft))*(bdrms-mean(bdrms))), data = hprice1)
gradethis_setup()
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(echo = FALSE)
```

## Square Footage, Bedrooms, and Home Prices 

Estimate the following models using the housing price dataset (`hprice1` - note that this is *not* the same dataset as the tutorial!).

1. $Price = \beta_0 + \beta_1Lot\text{ }Size + \beta_2Square\text{ }Feet + \beta_3Bedrooms + u \text{ : (hprice.lm1)}$
2. $Price = \beta_0 + \beta_1Lot\text{ }Size + \beta_2Square\text{ }Feet + \beta_3Bedrooms + \beta4Square\text{ }Feet \cdot Bedrooms + u \text{ : (hprice.lm2)}$
3. $Price = \beta_0 + \beta_1Lot\text{ }Size + \beta_2Square\text{ }Feet + \beta_3Bedrooms + \beta4(Square\text{ }Feet - \mu_{Sq.\text{ }Ft.}) \cdot (Bedrooms - \mu_{Bedrooms}) + u \text{ : (hprice.lm3)}$

Present all three sets of estimates in the same **text** table using `stargazer.`

```{r housePricePractice, exercise=TRUE, exercise.reveal_solution = FALSE}

```

```{r housePricePractice-solution}
hprice.lm1 <- lm(price ~ lotsize + sqrft + bdrms, data = hprice1)
hprice.lm2 <- lm(price ~ lotsize + sqrft + bdrms + sqrft*bdrms, data = hprice1)
hprice.lm3 <- lm(price ~ lotsize + sqrft + bdrms + I(scale(sqrft, scale = FALSE)*scale(sqrft, scale = FALSE)), data = hprice1)
stargazer(hprice.lm1, hprice.lm2, hprice.lm3, type = 'text')
```

```{r housePricePractice-check}
grade_code()
```

