---
title: "Chapter 14" 
subtitle: "Advanced Panel Data Methods"
description: >
  This tutorial introduces fixed effects and random effects panel methods. 
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = TRUE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE, fig.height= 9, fig.width=8)
```

## "Advanced" Panel Estimators

Types of panel estimators: 

  1.  First-difference estimator (already DiD)
  $$(y_{it} - y_{i,t-1}) = \beta_0 + \beta_1(x_{1it} - x_{1i,t-1}) + ... + \beta_k(x_{kit} - x_{ki,t-1}) + u_{it}$$
  2.  Between estimator
  $$\bar{y}_{i} = \beta_0 + \beta_1\bar{x}_{1i} + ... + \beta_k\bar{x}_{i} + \bar{u}_i$$
  3.  Within (time-demeaned) estimator
  $$(y_{it} - \bar{y}_{i}) = \beta_0 + \beta_1(x_{1it} - \bar{x}_{1i}) + ... + \beta_k(x_{kit} - \bar{x}_{ki}) + u_{it} - \bar{u}_i$$
  4.  Fixed effects estimator
  $$y_{it} = \beta_0 + \beta_1x_{1it} + ... + \beta_kx_{kit} + a_i + u_{it}$$
  5.  Random effects estimator
  $$(y_{it} - \theta\bar{y}_{i}) = \beta_0 + \beta_1(x_{1it} - \theta\bar{x}_{1i}) + ... + \beta_k(x_{kit} - \theta\bar{x}_{ki}) + u_{it} - \bar{u}_i$$
  $$0 < \theta < 1$$
  6.  Correlated random effects estimator
  
      We shall see...

## Fixed Effects and Within Estimator Exercises

Consider the following panel model: 

$$ln(Crime Rate) = \beta_0 + \beta_1ln(P_i[Arrest]) + \beta_2ln(P_i[Conviction]) + \beta_3ln(P_i[Prison]) + \beta_4ln(AveSentence_i) + \beta_5ln(PolicePC_i) + a_{i} + u_{it}$$

Estimate this model the following ways: 

  1. Calculate the OLS estimator with dummy variables for $county$ using $lm()$ ($crmrte.fe1$). 
  2. Construct the "within" transformation of all of the model variables ($variable.w$) and regress them on one another using $lm()$ ($crmrte.fe2$). 
  3. Calculate the "within" estimator using $plm$ with $model = "within"$ ($crmrte.fe3$).
  4. Calculate the "individual fixed effects" estimator using $plm$ with $effects = "individual"$  ($crmrte.fe4$).
  
Create a $stargazer$ text table of all four estimates, omitting the $county$ dummy coefficients from specification 1. 

```{r fe, exercise = TRUE}

```

```{r fe-hint}
crmrte.fe1 <- lm(formula, data)
crime4.p$lcrmrte.w <- Within(...)
crime4.p$lprbarr.w <- Within(...)
crime4.p$lprbconv.w <- Within(...)
crime4.p$lprbpris.w <- Within(...)
crime4.p$lavgsen.w <- Within(...)
crime4.p$lpolpc.w <- Within(...)
crmrte.fe2 <- lm(formula, data)
crmrte.fe3 <- plm(formula, data, model, effects)
crmrte.fe4 <- plm(formula, data, effects)
stargazer(list, type, omit)
```

```{r fe-solution}
crmrte.fe1 <- lm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + factor(county), 
                  data = crime4.p)
crime4.p$lcrmrte.w <- Within(crime4.p$lcrmrte)
crime4.p$lprbarr.w <- Within(crime4.p$lprbarr)
crime4.p$lprbconv.w <- Within(crime4.p$lprbconv)
crime4.p$lprbpris.w <- Within(crime4.p$lprbpris)
crime4.p$lavgsen.w <- Within(crime4.p$lavgsen)
crime4.p$lpolpc.w <- Within(crime4.p$lpolpc)
crmrte.fe2 <- lm(lcrmrte.w ~ lprbarr.w + lprbconv.w + lprbpris.w + lavgsen.w + lpolpc.w, 
                  data = crime4.p)
crmrte.fe3 <- plm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
                  data = crime4.p, model = "within", effects = "none")
crmrte.fe4 <- plm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
                  data = crime4.p, effects = "individual")
stargazer(crmrte.fe1, crmrte.fe2, crmrte.fe3, crmrte.fe4, type = 'text', omit = 'county')
```

```{r fe-check}
grade_code()
```

## Random Effects and Between Estimator Exercises

$$ln(Crime Rate) = \beta_0 + \beta_1ln(P_i[Arrest]) + \beta_2ln(P_i[Conviction]) + \beta_3ln(P_i[Prison]) + \beta_4ln(AveSentence_i) + \beta_5ln(PolicePC_i) + a_{i} + u_{it}$$

Estimate this model the following ways: 

  1. Construct the "between" transformation of all of the model variables ($variable.b$) and regress them on one another using $lm()$  ($crmrte.be$). 
  2. Calculate the random-effects estimator using $plm$ with $model = "random"$ ($crmrte.re1$).
  3. Construct $\theta$ using the formula $\theta = 1 - [\sigma_u^2/(\sigma_u^2+T\sigma_a^2)]^{1/2}$, and the corresponding transformations of the variables ($variable.r$) to calculate the estimate of the random-effects estimator "from scratch"(-ish) ($crmrte.re3$).

Create a $stargazer$ text table of these three estimates *and* the within estimator ($crmrte.fe3$). 

```{r re, exercise = TRUE}

```

```{r re-hint}
crime4.p$lcrmrte.b <- Between(...)
crime4.p$lprbarr.b <- Between(...)
crime4.p$lprbconv.b <- Between(...)
crime4.p$lprbpris.b <- Between(...)
crime4.p$lavgsen.b <- Between(...)
crime4.p$lpolpc.b <- Between(...)
crmrte.be1 <- lm(formula, data)
crmrte.re1 <- plm(formula, data, model, effects)
# The random-effects estimator in "plm" stores a value "ercomp$sigma2" that contains the components 
# of the errors: the first element is the variance of the idiosyncratic error, "u_it"; the second is 
# the variance of the individual-specific error, "a_i". 
theta <- 1 - sqrt(.../(...+...))
crime4.p$lcrmrte.r <- ...
crime4.p$lprbarr.r <- ...
crime4.p$lprbconv.r <- ...
crime4.p$lprbpris.r <- ...
crime4.p$lavgsen.r <- ...
crime4.p$lpolpc.r <- ...
crmrte.re2 <- lm(formula, data)
stargazer(list, type)
```

```{r re-solution}
crime4.p$lcrmrte.b <- Between(crime4.p$lcrmrte)
crime4.p$lprbarr.b <- Between(crime4.p$lprbarr)
crime4.p$lprbconv.b <- Between(crime4.p$lprbconv)
crime4.p$lprbpris.b <- Between(crime4.p$lprbpris)
crime4.p$lavgsen.b <- Between(crime4.p$lavgsen)
crime4.p$lpolpc.b <- Between(crime4.p$lpolpc)
crmrte.be1 <- lm(lcrmrte.b ~ lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                 data = crime4.p)
crmrte.re1 <- plm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
                  data = crime4.p, model = "random", effects = "none")
theta <- 1 - sqrt(crmrte.re1$ercomp$sigma2[1]/(crmrte.re1$ercomp$sigma2[1]+7*crmrte.re1$ercomp$sigma2[2]))
crime4.p$lcrmrte.r <- crime4.p$lcrmrte - theta*Between(crime4.p$lcrmrte)
crime4.p$lprbarr.r <- crime4.p$lprbarr - theta*Between(crime4.p$lprbarr)
crime4.p$lprbconv.r <- crime4.p$lprbconv - theta*Between(crime4.p$lprbconv)
crime4.p$lprbpris.r <- crime4.p$lprbpris - theta*Between(crime4.p$lprbpris)
crime4.p$lavgsen.r <- crime4.p$lavgsen - theta*Between(crime4.p$lavgsen)
crime4.p$lpolpc.r <- crime4.p$lpolpc - theta*Between(crime4.p$lpolpc)
crmrte.re2 <- lm(lcrmrte.r ~ lprbarr.r + lprbconv.r + lprbpris.r + lavgsen.r + lpolpc.r, 
                  data = crime4.p)
stargazer(crmrte.be1, crmrte.re1, crmrte.re2, crmrte.fe3, type = 'text')
```

```{r re-check}
grade_code()
```

## A Short Aside on Random Effects - "Feasible Generalized Least Squares" (FGLS)

Technically, the model we estimate 
$$ln(Crime Rate) = \beta_0 + \beta_1ln(P_i[Arrest]) + \beta_2ln(P_i[Conviction]) + \beta_3ln(P_i[Prison]) + \beta_4ln(AveSentence_i) + \beta_5ln(PolicePC_i) + u_{it}$$

$$u_{it} = a_{i} + \eta_{it}$$

  1. Estimate pooled OLS
  2. Regress residuals on fixed-effects
  3. Get $\theta$ by finding the proportion of the standard error of $u_{it}$ that comes from $\eta_{it}$ 
  4. Repeat until reach convergence (estimates stop changing)


## Correlated Random Effects

$$ln(Crime Rate) = \beta_0 + \beta_1ln(P_i[Arrest]) + \beta_2ln(P_i[Conviction]) + \beta_3ln(P_i[Prison]) + \beta_4ln(AveSentence_i) + \beta_5ln(PolicePC_i) + a_{i} + u_{it}$$

Estimate this model the following ways: 

  1. Use a random-effects estimator, adding the *between transforms* ($variable.b$) of all of the time-varying variables as additional controls ($crmrte.cre1$). 
  2. Use a random-effects estimator with between-effects controls, adding region ($west$ and $central$) and $urban$ as controls ($crmrte.cre2$).
  3. Use a fixed-effects estimator, *without* between-effects controls, adding region ($west$ and $central$) and $urban$ as controls ($crmrte.fe5$).

Create a $stargazer$ text table of these estimates and the fixed-effects estimator ($crmrte.fe4$). 

```{r cre, exercise = TRUE}

```

```{r cre-hint}
crmrte.cre1 <- plm(formula, data, model, effects)
crmrte.cre2 <- plm(formula, data, model, effects)
crmrte.fe5 <- plm(formula, data, effects)
stargazer(list, type)
```

```{r cre-solution}
crmrte.cre1 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + 
                    lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.cre2 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central +
                    urban + lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.fe5 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central + 
                   urban, data = crime4.p, effects = "individual")
stargazer(crmrte.cre1, crmrte.cre2, crmrte.fe5, crmrte.fe4, type = 'text')
```

```{r cre-check}
grade_code()
```

## Clustered Sampling, Hierarchical Models, and Other Uses of Panel Methods

