---
title: "Chapter 6"
subtitle: "Multiple Regression Analysis - Scaling and Elasticities"
description: "This tutorial practices standardizing variables, elasticities, and log transformations."
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = FALSE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```

## Crime and House Prices

### Standardized Coefficients

1. Estimate the following model using the `hprice2` dataset as `hprice.lm1`: 
$$Price = \beta_0 + \beta_1NO2 + \beta_2Crime + \beta_3Rooms + \beta4Employer\text{ }Distance + \beta_5Student\text{-}Teacher\text{ }Ratio + u \text{ : (hprice.lm1)}$$ 
2. Reestimate the same model *with and without a constant* using the `scale` function in your formula argument to estimate the standardized coefficients (`hprice.lm2` and `hprice.lm3`). 

```{r housePriceBetas, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r housePriceBetas-solution}
library(stargazer)
hprice2 <- wooldridge::hprice2
hprice.lm1 <- lm(price ~ nox + crime + rooms + dist + stratio, data = hprice2)
hprice.lm2 <- lm(scale(price) ~ scale(nox) + scale(crime) + scale(rooms) + scale(dist) + scale(stratio), data = hprice2)
hprice.lm3 <- lm(scale(price) ~ 0 + scale(nox) + scale(crime) + scale(rooms) + scale(dist) + scale(stratio), data = hprice2)
stargazer(hprice.lm1, hprice.lm2, hprice.lm3, type = 'text')
```

```{r housePriceBetas-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Elasticities

Calculate the elasticities for the model estimated in (1) and present the estimates in the same text table using `stargazer()`.

```{r housePriceEYEX, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r housePriceEYEX-setup}
library(stargazer)
hprice2 <- wooldridge::hprice2
hprice.lm1 <- lm(price ~ nox + crime + rooms + dist + stratio, data = hprice2)
hprice.lm2 <- lm(scale(price) ~ scale(nox) + scale(crime) + scale(rooms) + scale(dist) + scale(stratio), data = hprice2)
hprice.lm3 <- lm(scale(price) ~ 0 + scale(nox) + scale(crime) + scale(rooms) + scale(dist) + scale(stratio), data = hprice2)
```

```{r housePriceEYEX-hint}
# `stargazer()` has options to replace the coefficients (`coef = list()`) or standard errors (`se = list()`) with marginal effects or robust standard errors (see chapter 8).
# The `data.frame` that `summary.marginaleffects()` stores the coefficients and standard errors in columns named `estimate` and `std.error`. 
```

```{r housePriceEYEX-solution}
library(marginaleffects)
hprice.eyex1 <- marginaleffects(hprice.lm1, slope = 'eyex') |> 
  summary()
stargazer(hprice.lm1,                              # Tell it the structure of the original model
          coef = list(hprice.eyex1[['estimate']]), # Tell it where to find the elasticities
          se = list(hprice.eyex1[['std.error']]),  # Tell it where to find the se's
          type = 'text')                           # Remember to specify type
```

```{r housePriceEYEX-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Log Transofrmation

Estimate the following model and print a `summary()` of the result.  
$$ln(Price) = \beta_0 + \beta_1ln(NO2) + \beta_2ln(Crime) + \beta_3Rooms + \beta_4Rooms^2 + \beta5Employer\text{ }Distance + \beta_5Student\text{-}Teacher\text{ }Ratio + u \text{ : (lnhprice.lm1)}$$ 
```{r lnHousePrice, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r lnHousePrice-setup}
hprice2 <- wooldridge::hprice2
```

```{r lnHousePrice-solution}
lnhprice.lm1 <- lm(log(price) ~ log(nox) + log(crime) + rooms + I(rooms^2) + log(dist) + log(stratio), data = hprice2)
summary(lnhprice.lm1)
```

```{r lnHousePrice-hint}
# Remember that in order to use math operations (*, /, ^, etc.) within a formula, you need to wrap these terms within the I() function (but not with functions like `log()` or `sqrt()`. 
```

```{r lnHousePrice-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```
