---
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
title: "Appendix C"
subtitle: |
  | Fundamentals of Mathematical Statistics
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
always_allow_html: yes
description: "This tutorial practices the concepts of the laws of large numbers and central limit theorem."
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis_setup()
tutorial_options(exercise.reveal_solution = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

## Simple Regression Simulation

Let $X$ be a random variable that follows a exponential distribution with rate parameter $\lambda = 1$. 
Let $Y$ be a random variable that depends on $X$ according to the known linear structural population model $Y = \beta_0 + \beta_1X + u$, where: 
  
  + $\beta_0 = 2$
  + $\beta_1 = -0.5$
  + $u \sim N(0,1)$

Create a loop that does the following: 

1. Simulate samples of sizes 1-1000 for $X$, $u$, and $Y$.
2. Estimates the equation for Y to obtain $\hat\beta_0$ and $\hat\beta_1$ for each sample size.
3. Show the convergence of the sample estimate of $\beta_1$ towards the known population value of -1 for sample sizes 1 through 1000 by plotting the squared deviations from the true value.
4. Print the 'head()` of your simulation output.

```{r llnPractice, exercise=TRUE, exercise.eval=TRUE}
set.seed(8675309)
X <- NULL 
u <- NULL
Y <- NULL
b1 <- NULL


```

```{r llnPractice-hint-1}
# Don't forget to set the random seed for reproducibility.
# Create "NULL" objects for X, u, Y, and b1. 
# Define a for-loop over different sample sizes. 
```

```{r llnPractice-hint-2}
# The elements of X are random exponential draws [`rexp()`] with rate parameter 1.
# The elements of u are random normal draws [`rnorm()`] with mean 0 and sd 1. 
# The elements of Y are drawn from the known relationship of Y with X (b0 + b1X + u)
# Use the "known" values of b0 and b1 (2 and -0.5), with the random u's to generate Y. 
```

```{r llnPractice-hint-3}
# The `lm()` function estimates the linear regression.
# The named `coefficients` value from the `lm()` output extracts the vector of betas
# The slope beta (b1) is the 2nd element of the coefficient vector: 
#     b1 = lm(...)$coefficients[2]
```

```{r llnPractice-solution}
set.seed(8675309)
X <- NULL 
u <- NULL
Y <- NULL
b1 <- NULL
for(i in 1:1000) {
  X = rexp(n = i, rate = 1)
  u = rnorm(n = i, mean = 0, sd = 1)
  Y = 2 - 0.5*X + u
  b1[i] <- lm(Y ~ X)$coefficients[2]
}
df <- data.frame(n = c(1:length(b1)), b1 = b1, v1 = (b1-(-0.5))^2)
plot(c(1:length(X)), df$v1, type = 'l', 
     xlab = "Sample Size", ylab = "Squared Deviation", 
     ylim = c(0,2))

```

```{r llnPractice-check}
grade_this(if (!inherits(.result, "data.frame")) {
    fail("Your class of your answer should be 'data.frame'.")}
  if(length(.result) != 3) {
    fail("Your table should have a length of 3.") } 
  if(round(.result[2,2], 4) != 2.0112) {
    fail("For n=2 b1 should equal about 2.0112.") } 
  pass()
)
```

## Regression CLT Practice

Using the same population parameters and variables as the previous exercise, create a loop that does the following: 

1. Simulate 1000 *distributions* of $\hat\beta_1$ with 1000 values estimates each from samples $X$, $u$, and $Y$ ranging from 1-1000. 
2. Estimates the equation for Y to obtain $\hat\beta_0$ and $\hat\beta_1$ for each sample size.
3. Plot the convergence of the distribution of $\hat\beta_1$ towards a normal distribution. 
4. Print the `head()` of the matrix of values for $\beta_1$. 

```{r cltPractice, exercise=TRUE, exercise.eval=TRUE}
set.seed(8675309)


```

```{r cltPractice-hint}
# This is similar to the LLN example, but with a double loop (sample sizes and resamples).
```

```{r cltPractice-solution}
set.seed(8675309)
X <- NULL 
u <- NULL
Y <- NULL
b1 <- matrix(NA, nrow = 1000, ncol = 100)
for(i in 1:100) {
  for(j in 1:1000) {
    X <- rexp(n = i, rate = 1) 
    u <- rnorm(n = i, mean = 0, sd = 1)
    Y = 2 - 0.5*X + u
    b1[j, i] = lm(Y ~ X)$coefficients[2]
  }
}
plot(density(b1[,2]), xlim = c(-2, 1), ylim = c(0, 4))
lines(density(b1[,10]))
lines(density(b1[,30]))
lines(density(b1[,100]))
head(b1)
```

```{r cltPractice-check}
grade_this(if (!inherits(.result, c("matrix", "array"))) {
    fail("Your class of your answer should be c('matrix', 'array').")}
  if(length(.result) != 600) {
    fail("Your table should have a length of 600.") } 
  if(round(.result[2,2], 4) != 8.5484) {
    fail("For n=2 b1 should equal about 8.5484") } 
  pass()
)
```
