---
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
title:  "Chapter 1" 
subtitle:  |
  | 1. Introduction
  | 2. Installing R & RStudio
  | 3. Nature of Econometrics & Data
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
always_allow_html: yes
description: >
  This tutorial introduces students to the course and demonstrates the basic setup for R and RStudio.
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 14pt}
</style>
```
```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis_setup()
tutorial_options(exercise.eval = FALSE, exercise.checker = gradethis_exercise_checker)
knitr::opts_chunk$set(echo = FALSE, tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Syllabus Information

### General Information

-   Office Hours: MWF 11am-12:30pm; TR 8:30-10am; liberally by
    appointment: [book me
    here](https://calendar.google.com/calendar/selfsched?sstoken=UUFLaDRHQXJVeXZHfGRlZmF1bHR8YjU1MzlmMzk2YjBlMWEzMzUzNTUzZDI0MjVhYzBiMTMA)!
-   Research Paper Deadlines:
-   Class Preparation, Participation, and Discussion
    -   The class participant's engagement actively contributes to
        general student learning -- particularly important with a very
        small class.

### What is Econometrics?

Simply put, econometrics seeks to provide valid answers to the question,
"How much?"\
- The quantitative analysis of economic data - Helps to quantify
relationships

We will use statistical analysis to test economic hypotheses.
Econometrics is quantitatively rigorous, but also requires economic
intuition. Analyzing data in a modern environment also requires
significant computing skills.

### Learning Objectives

The successful student will be able to:

1.  analyze economic relationships using regression analysis;
2.  interpret regression results;
3.  test the validity of the assumptions for regression models;
4.  conduct empirical economic research;
5.  perform statistical-computing operations with appropriate packages.

### Class Resources

Required:\
1. Wooldridge, J. 2020/2016. Introductory Econometrics, 7th Edition
(Recommended) or [5th
Edition](https://economics.ut.ac.ir/documents/3030266/14100645/Jeffrey_M._Wooldridge_Introductory_Econometrics_A_Modern_Approach__2012.pdf),
Cengage, Boston.\
2. Heiss, Florian. 2016. [Using R for Introductory
Econometrics](http://www.urfie.net/read/index.html). 3. Bang, J. 2023.
[`metricsToTheFace`: `learnr` Tutorials for *Introductory Econometrics*
using Wooldridge
Examples](https://github.com/bangecon/metricsToTheFace), Version 0.0.3.
4. Blackboard, Inc. 2021. [Blackboard](https://blackboard.sau.edu). 5. R
Core Team. 2021. [R: A Language and Environment for Statistical
Computing](https://cran.r-project.org/). R Foundation for Statistical
Computing. 6. RStudio Team. 2020. [RStudio: Integrated Development for
R](https://www.rstudio.com/products/rstudio/download/). Posit, Boston,
MA.

Recommended:\
1. Illowsky, Barbara and Susan Dean. 2013. [Introductory
Statistics](https://openstax.org/details/books/introductory-statistics).
Openstax. 2. Katchova, Ani, Econometrics Course, Econometrics Academy
MOOC. 3. Khan, Sal. Statistics and Probability, Khan Academy.

### Coursework and Grading

| Assignment Type                             | Point Total |
|---------------------------------------------|-------------|
| Paper Sections/Drafts (4 \@ 50 points each) | 200 points  |
| Peer Referee Report                         | 50 points   |
| Pilot Analysis Presentation                 | 50 points   |
| Final Draft of Research Paper               | 250 points  |
| Assessments                                 | 50 points   |
| Participation                               | 200 points  |
| Final Exam                                  | 200 points  |

<br> <br>

| Letter Grade | Point Range      |
|--------------|------------------|
| A+/A/A-      | 900-1,000 points |
| B+/B/B-      | 800-899.9 points |
| C+/C/C-      | 700-799.9 points |
| D+/D/D-      | 600-699.9 points |
| F            | \< 600 points    |

### Class Structure

This will be a "flipped" class of sorts. I will expect you to work
through the main tutorial for each topic before class, record your
progress on the exercises for that tutorial, and come to class ready to
give a brief presentation of your work if I call on you.

At the start of class, I will ask for a volunteer to present the main
tutorial (15-20 minutes?). If no one volunteers, I will select one
person at random. Coming to class unprepared to present will affect your
participation grade if you are selected, as will absence from class
(regardless of whether you are selected to present). After that, we will
work the practice tutorial for that topic together-ish.

You will start the semester with 2 tokens that you can use for various
forms of leniency, including:\
1. Forgiving a class absence; 2. Forgiving being unprepared to present
on a day that no one volunteers and you are selected; 3. Replace a
missed assessment score; 4. A one-class extension of a paper deadline;
5. Revising a paper submission.

## Getting Started with R

### Why R?

R helps you talk to a computer and help solve big problems. Employers
use R (or other similar languages like Python or Julia) for statistics
and data analysis. We will use R in this class to develop basic
proficiency with a computing language used in industry

### Installing R

```{r installR, echo=FALSE}
question("Where do you download R?",
    answer("[www.rstudio.com/download](www.rstudio.com/download)"),
    answer("[cloud.r-project.org](http://cloud.r-project.org)", correct = TRUE, message = "You can also download R from [cran.r-project.org](http://cran.r-project.org)"),
    answer("[www.r-project.org](www.r-project.org)", message = "Good try, but not exactly. www.r-project.org doesn't provide a download link, but it does provide a link to one the websites above."),
    answer("[www.r.com](www.r.com)"),
    allow_retry = TRUE
  )
```

### Updating R

```{r updaterR, echo=FALSE}
question("How often should you update R?",
    answer("Everytime you use it", message = "This will be too often unless you use R very rarely!"),
    answer("About once a year", correct = TRUE, "A new version of R is released about once a year. Update sooner if you encounter a bug that you cannot explain."),
    answer("Never", message = "A new version of R is released about once a year. I'll assume that you are using the newest version of R, which will be the fastest version with the fewest unexpected behaviors." ),
    allow_retry = TRUE
  )
```

## RStudio

### What is RStudio?

```{r rStudio, echo=FALSE}
  question("What is the RStudio IDE?",
    answer("An application that makes it easier to use R.", correct = TRUE, message = "RStudio organizes your workspace and makes it easier to write, use, debug, and save R code. I highly recommend it and will rely on the RStudio IDE in several tutorials."),
    answer("An application that let's you use R without writing any code", message = "And thank goodness! Code provides a reproducible record of your work, which is essential for data science."),
    answer("A spreadsheet program like Microsoft Excel."),
    answer("Another name for R", message = "R and RStudio are two separate things. R is a language, like English. Think of RStudio as a program that helps you use the language, kind of like how a word processing program helps you write in English."),
    allow_retry = TRUE
  )
```

### Why RStudio?

RStudio is an Integrated Development Environment (IDE) for R. What does
that mean? Well, if you think of R as a language, which it is, you can
think of RStudio as a program that helps you write and work in the
language. RStudio makes programming in R much easier and I suggest that
you use it!

### Where do I get RStudio?

```{r installRStudio, echo=FALSE}
question("Where do you download RStudio?",
    answer("[www.rstudio.com/download](www.rstudio.com/download)", correct = TRUE, message = "For these tutorials, download and install the RStudio Desktop Open Source License."),
    answer("[cloud.r-project.org](http://cloud.r-project.org)", message = "This is where you download R, not RStudio."),
    answer("[www.r-project.org](www.r-project.org)"),
    answer("[cran.rstudio.org](http://cran.rstudio.org)"),
    allow_retry = TRUE
  )
```

### Do I still need R?

```{r installR_RStudio, echo=FALSE}
  question("Do you need to install R if you already have RStudio?",
    answer("Yes.", correct = TRUE),
    answer("No.", message = "R does not come with RStudio; you need to install R separately.")
  )
```

## R Packages

### How do I install packages?

```{r install_packages, echo = FALSE}
  question("What command do you use to install packages?",
    answer("`library()`", message = "We will see what `library()` does later."),
    answer("`install.packages()`", correct = TRUE),
    answer("`install_packages()`"),
    answer("There is no command. You must visit [cran.r-project.org](http://cran.r-project.org) and download packages manually.", message = "R makes it easy to download packages. You connect to the internet and then run one of the commands above."),
    allow_retry = TRUE
  )
```

### How often do I install a package?

```{r reinstall_packages, echo = FALSE}
  question("How often do you need to install a package on your computer?",
    answer("Every time you restart R"),
    answer("Every time you restart your computer"),
    answer("Once.", correct = TRUE, message = "Afterwards, R can find it on your hard drive and load it using 'library()' or 'require()' as needed."),
    answer("Never, as long as you are connected to the internet.", message = "This could be true if you are using R over a cloud service. However, if you are using R locally on your own computer, you will need to install each package that you use on your computer."),
    allow_retry = TRUE
  )
```

## Economics and the Scientific Method

1.  Ask a Question
2.  Form a Hypothesis
3.  Review Existing Research
4.  Collect Data
5.  Analyze the Data

-   Test *Causal* Hypotheses
-   Forecast/Predict

6.  Form a Conclusion

### Example 1: Effects of Training on Productivity

$$wage=f(educ,exper,training; u)$$ Where:\
- $wage$ is the worker's hourly wage, - $educ$ is a measure of
schooling, - $exper$ is a measure of workforce experience,\
- $training$ is a measure of time spent training, and - $u$ is random
noise.

Assuming we can approximate $f(\cdot)$ with a linear function, we could
rewrite this as:

$$wage=\beta_0 + \beta_1educ + \beta_2exper + \beta_3training + u$$

An econometric model consists of a systematic part, $X\beta$, and a
random error, $u$

### Example 2: Effects of Sentencing on Crime

$$crime=f(crimewage, marketwage, otherincome, p_{caught}, p_{convicted|caught}, \hat{sentence}, age; u)$$

Where:\
- $crime$ is the number of hours spent in illegal activities, -
$crimewage$ is the "wage" per hour for illegal work, - $marketwage$ is
the wage for legal employment, - $otherincome$ is income from other
sources, - $p_{caught}$ is the probability of getting caught, -
$p_{convicted|caught}$ is the probability of conviction if caught, -
$\hat{sentence}$ is the expected sentence if convicted, and - $age$ is
the person's age.

Some variables might not be observed (crime wage); others might only be
observe by proxy or as an average ($p_{caught}$, $p_{convicted|caught}$,
$\hat{sentence}$)

$$crime=\beta_0 + \beta_1marketwage + \beta_2otherinc + \beta_3freqarr + \beta_4freqconv + \beta_5avesen + \beta_6age + u$$

## Data: Sources and Types

-   Experimental versus Observational Data
-   Cross-section
-   Time-series data
-   Pooled cross sections
-   Panel/longitudinal data

## Causality

How could we determine whether spending on education *causes* better
performance?

Hold all else equal: *ceteris paribus*

### Problems:

1.  Confounding (omitted) variables\
    Solution: Add those variables to the model.
2.  Non-random sample selection\
    Solution: Tobit (Chapter 17; Tobin, 1958); Heckit (Chapter 17;
    Heckman, 1976)
3.  Non-random treatment assignment\
    Solution: Instrumental Variables (Chapter 15); Matching treatment
    and controls (Rosenbaum & Rubin, 1983); Covariate balancing (Horvitz
    & Thompson, 1952; Robins et al., 2000)

## Some R Basics

### Classes of Objects

+-----------------+-----------------+-------------------------------+
| Class           | Description     | Example                       |
+=================+=================+===============================+
| numeric,        | Numbers         | 1, 2.5, pi(), exp()           |
| integer         |                 |                               |
+-----------------+-----------------+-------------------------------+
| character       | Letters &       | "John Doe", "Alaska",         |
|                 | numbers         | c("Red", "Blue", "Green")     |
+-----------------+-----------------+-------------------------------+
| factor          | Labelled        |                               |
|                 | categories      |                               |
+-----------------+-----------------+-------------------------------+
| logical         | "TRUE" or       | `is.na()` result; `2==2`      |
|                 | "FALSE"         |                               |
+-----------------+-----------------+-------------------------------+
| vector, matrix  | Sequence of     | `c(1,2,3,4)`;\                |
|                 | numbers,\       | `matrix(1:4, nrow = 2)`       |
|                 | Sequence of     |                               |
|                 | columns         |                               |
+-----------------+-----------------+-------------------------------+
| list            | Multiple        |                               |
|                 | objects of      |                               |
|                 | different sizes |                               |
|                 | or types        |                               |
+-----------------+-----------------+-------------------------------+
| data.frame      | List of         | `wooldridge::wage1`           |
|                 | variables with  |                               |
|                 | same lengths    |                               |
+-----------------+-----------------+-------------------------------+
| formula         | Linear formula  | `y ~ x1 + x2`                 |
|                 | *excluding*     |                               |
|                 | parameters      |                               |
+-----------------+-----------------+-------------------------------+
| lm              | Linear model    | `                             |
|                 | result          | y.lm <- lm(y ~ x, data = df)` |
+-----------------+-----------------+-------------------------------+

### Common Functions for Econometrics

+------------+-----------+-----------+-----------+--------------------+
| Function   | Package   | Arguments | De        | Example            |
|            |           |           | scription |                    |
+============+===========+===========+===========+====================+
| `help()`,  | utils     | topic     | Help with | `help(lm)`, `?lm`  |
| `?`        |           |           | a         |                    |
|            |           |           | function  |                    |
+------------+-----------+-----------+-----------+--------------------+
| `help.s    | utils     | pattern   | Help with | `help.search("     |
| earch()`,\ |           |           | a topic   | linear models")`,\ |
| `??`       |           |           |           | `                  |
|            |           |           |           | ??"linear models"` |
+------------+-----------+-----------+-----------+--------------------+
| `          | base      | object    | Quickly\  | `summary(df)`      |
| summary()` |           |           | s         |                    |
|            |           |           | ummarize\ |                    |
|            |           |           | da        |                    |
|            |           |           | ta/models |                    |
+------------+-----------+-----------+-----------+--------------------+
| `tbl_      | gtsummary | data,     | Summary\  | `tbl_summary(df,`\ |
| summary()` |           | statistic | tables    | `s                 |
|            |           |           |           | tatistic = list(`\ |
|            |           |           |           | `all_ continuous~  |
|            |           |           |           | "{mean} ({sd})"))` |
+------------+-----------+-----------+-----------+--------------------+
| `plot()`   | graphics  | x, y      | Sc        | `plot(df$x, df$y)` |
|            |           |           | atterplot |                    |
+------------+-----------+-----------+-----------+--------------------+
| `ggplot()` | ggplot2   | data,     | Begin gg  | `ggplot(           |
|            |           | mapping   |           | df, aes(x, y)) +`\ |
|            |           |           |           | `geom_point()`     |
+------------+-----------+-----------+-----------+--------------------+
| `lm()`     | stats     | formula,  | Linear    | `lm(y ~ x1         |
|            |           | data      | model     |  + x2, data = df)` |
+------------+-----------+-----------+-----------+--------------------+
| `st        | stargazer | model,    | Re        | `                  |
| argazer()` |           | type      | gression\ | stargazer(lm1, lm2 |
|            |           |           | tables    | , type = "text",`\ |
|            |           |           |           | `f                 |
|            |           |           |           | ile = "file.txt")` |
+------------+-----------+-----------+-----------+--------------------+
