---
output: slidy_presentation
runtime: shiny_prerendered
title: "Chapter 15 \n\n Alternatives to Instrumental Variables"
author: "James Bang"
description: >
  This tutorial introduces matching and balancing methods. 
---

```{r setup, include=FALSE}
library(tidyverse)
library(haven)
library(stargazer)
library(MatchIt)
library(cem)
library(estimatr)
library(gtsummary)

gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```

## Simple Mean Comparison

Question: Were first-class passengers more likely to survive on the Titanic? 

```{r simpleMeans, echo = TRUE}
titanic <- read_dta("https://raw.github.com/scunning1975/mixtape/master/titanic.dta") 
titanic$FirstClass <- (titanic$class == 1)
titanic$age <- (titanic$age == 1)
titanic$sex <- (titanic$sex == 0)
titanic$survived <- (titanic$survived == 1)
colnames(titanic) <- c("Class", "Adult", "Female", "Survived", "FirstClass")
add_p(tbl_summary(titanic[,2:5], by = FirstClass))
```

## Controlled Least-Squares Comparison

Using OLS and controlling for age (adult versus child) and gender tells a slightly different story: 

```{r controlledOLS, echo = TRUE, results = 'asis'}
titanic.lm <- lm(Survived ~ FirstClass + Adult + Female, data = titanic)
stargazer(titanic.lm, type = 'html')
```

## Matching: A Brief Overview of Methods

1. Propensity Scores
2. Distance Matching
    - Euclidian Distance
    - Mahalanobis Distance
3. Exact Matching
4. Coarsened Exact Matching

Most matching mixes a combination of these. 

Sometimes, we even use a controlled regression *after* matching on some of the observables.

## Matched Mean Comparison

Matching (exactly) according to age (adult versus child) and gender tells an even more different story: 

```{r matchedMeans, echo = TRUE}
classMatch <- matchit(FirstClass ~ Adult + Female, data = titanic, method = 'nearest')
titanic.matchit  <- match.data(classMatch)
add_p(tbl_summary(titanic.matchit[,2:5], by = FirstClass))
```

