---
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
title: "Chapter 19"
subtitle: |
  | Conducting & Presenting Research: Tables
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
description: >
  This tutorial practices tools for creating attractive tables and graphs.
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis::gradethis_setup()
tutorial_options(exercise.reveal_solution = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

## Summarizing Data

### Summarizing an Entire Data Frame

Calculate the summary statistics for the *salary*, *sales*, *roe*, and *ros* variables. 

```{r multipleSummaryPractice, exercise=TRUE, exercise.eval=TRUE}

```

```{r multipleSummaryPractice-hint}
# Don't forget to load your data!
```

```{r multipleSummaryPractice-solution}
ceosal1 <- wooldridge::ceosal1
summary(ceosal1[ , c('salary', 'sales', 'roe', 'ros')])
```

```{r multipleSummaryPractice-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Summaries with Categorical Variables

Using the `ceosal1` dataset, create a new variable named `industry` defined as follows: 

$$Industry = \begin{cases} \begin{aligned}
&1 &\text{ if } &indus = 1 \\
&2 &\text{ if } &finance = 1 \\
&3 &\text{ if } &consprod = 1 \\
&4 &\text{ if } &utility = 1 \\
\end{aligned} \end{cases}$$

Summarize the variable you create. 

```{r summaryFactors, exercise=TRUE, exercise.eval=FALSE}
# Create a factor variable for the industry


# Summarize it with `summary()`

```

```{r summaryFactors-setup}
ceosal1 <- wooldridge::ceosal1
```

```{r summaryFactors-hint}
# Don't forget to attach the data or include the data frame when you call each variable.
```

```{r summaryFactors-solution}
A <- cbind(ceosal1$indus, ceosal1$finance, ceosal1$consprod, ceosal1$utility) 
f <- c(1:ncol(A))
ceosal1$industry <- factor(A %*% f, labels = c('Industrial', 'Finance', 'Consumer Products', 'Utilities'))
summary(ceosal1$industry)
```

```{r summaryFactors-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Summary by Categorical Group

Create a `gtsummary::` table for `salary`, `sales`, `roe`, and `ros` *by industry*. 

```{r gtsummaryPractice, exercise=TRUE, exercise.eval=TRUE}

```

```{r gtsummaryPractice-setup}
ceosal1 <- wooldridge::ceosal1
A <- cbind(ceosal1$indus, ceosal1$finance, ceosal1$consprod, ceosal1$utility) 
f <- c(1:ncol(A))
ceosal1$industry <- factor(A %*% f, labels = c('Industrial', 'Finance', 'Consumer Products', 'Utilities'))
```

```{r gtsummaryPractice-hint}
# Don't forget to load your libraries!
# Use square column-indexing brackets with "c('var1', ...)" 
# Don't forget to include your group variable in your indexing!
# Use the `by` option to create separate columns for each industry. 
```

```{r gtsummaryPractice-solution}
library(gtsummary)
tbl_summary(ceosal1[, c('salary', 'sales', 'roe', 'ros', 'industry')], by = industry, 
            statistic = list(all_continuous() ~ '{mean} ({sd})'))
```

```{r gtsummaryPractice-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Summarizing Relationships between Variables

Create a `corrgram` matrix with only correlation coefficients among `salary`, `sales`, `roe`, and `ros`. 

```{r correlationPractice, exercise=TRUE, exercise.eval=TRUE}

```

```{r correlationPractice-setup}
ceosal1 <- wooldridge::ceosal1
```

```{r correlationPractice-hint}
# Use square column-indexing brackets with "c('var1', ...)" 
```

```{r correlationPractice-solution}
library(corrgram)
corrgram(ceosal1[ , c('salary', 'sales', 'roe', 'ros')], panel = panel.cor)
```

```{r correlationPractice-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Other `corrgram` Options

Re-draw the previous `corrgram` with correlation coefficients along the lower panel and scatterplots on the upper panel.

```{r corrgramPractice, exercise = TRUE}

```

```{r corrgramPractice-setup}
library(corrgram)
ceosal1 <- wooldridge::ceosal1
```
  
```{r corrgramPractice-hint}

```

```{r corrgramPractice-solution}
corrgram(ceosal1[ , c('salary', 'sales', 'roe', 'ros')], upper.panel = panel.pts, lower.panel = panel.cor)
```

```{r corrgramPractice-check}
grade_this(if (identical(.result, .solution)) {
    pass("You can also set the `col.regions` option to change the color palette for the color contrasts.")
  } {
  fail(random_encouragement)
})
```

## Factor Variable Tables

The `ceosal1` data does not have factors other than industry. For practice, we can use `cut()` to make a categorical variable according to the quartiles of `salary`. Do this, and create a crosstabulation between `industry` and the quartile ranks of `salary`. Use `gtsummary::tbl_cross()` to make your table look nice. 

```{r gtcross, exercise=TRUE, exercise.eval=TRUE}


```

```{r gtcross-setup}
library(gtsummary)
ceosal1 <- wooldridge::ceosal1
```

```{r gtcross-hint}
# The margin option helps you specify which variable to condition the proportions on. 
```

```{r gtcross-solution}
tbl_cross(ceosal1, row = salary.q4, col = industry, percent = 'cell')
```

```{r gtcross-check}
grade_this(if (identical(.result, .solution)) {
    pass("You can also set the `col.regions` option to change the color palette for the color contrasts.")
  } {
  fail(random_encouragement)
})
```
