---
title: "Chapter 4"
subtitle: "Multiple Regression Analysis - Inference"
description: "This tutorial introduces inference for single-parameter tests following multiple regression."
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
gradethis::gradethis_setup()
tutorial_options(exercise.reveal_solution = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

## Population Density and Crime 

1. Estimate the following model using the `crime2` dataset.
$$ln(CrimeRate) = \beta_0 + \beta_1ln(LawExpendPC) + \beta_2ln(IncomePC) + \beta_3ln(PopulationDensity) + \beta_4West + \beta5Northeast + \beta_6South + u$$
2. Construct a 99% confidence interval for the law-enforcement expenditure elasticity of crime rates using `confint`.

```{r crime, exercise=TRUE}

```

```{r crime-solution, exercise.reveal_solution = FALSE}
crime2 <- wooldridge::crime2
crime.lm1 <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden) + west + nrtheast + south, data = crime2)
confint(crime.lm1, "log(lawexpc)", level = 0.99)
```

```{r crime-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Skipping Class (Also a Crime!)

### Estimation

Use the `colGPA` to estimate the following model:  
$$CollegeGPA = \beta_0 + \beta_1HSGPA + \beta_2ACT + \beta_3Skipped + u,$$ 
where `skipped` measures the average number of classes a student missed each week. 

Print a `summary()` of your results.

```{r gpaSkip, exercise = TRUE}

```

```{r gpaSkip-hint}

```

```{r gpaSkip-solution}
gpa1 <- wooldridge::gpa1
colGPA.lm2 <- lm(colGPA ~ hsGPA + ACT + skipped, data = gpa1)
summary(colGPA.lm2)
```

```{r gpaSkip-check}
grade_this({
  if (identical(.result$call, .solution$call)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Testing

Test the hypothesis that skipping more classes decreases your college GPA at the $\alpha = 0.01$ level of significance.  
Report the test statistic, degrees of freedom, t-critical value, and p-value for the test as a `c()` vector.  

```{r skipTest, exercise = TRUE}

```

```{r skipTest-hint-1}
# The test statistic is stored as part of the `summary.lm()` object in the `coefficients` table in the row named for the variable and the column named `t value`. 
```

```{r skipTest-hint-2}
# The degrees of freedom are stored in the `lm` object itself under the name `df.residual`. 
```

```{r skipTest-hint-3}
# The t-critical value is the quantile of the t-distribution, `qt()`, corresponding to placing probability `alpha` in the left tail (left-tailed test, lower.tail = TRUE, default), `alpha` in the right tail (right-tailed test, lower.tail = FALSE), or `alpha/2` in each of the left and right tails. 
```

```{r skipTest-setup}
gpa1 <- wooldridge::gpa1
colGPA.lm2 <- lm(colGPA ~ hsGPA + ACT + skipped, data = gpa1) 
```

```{r skipTest-solution}
tstat <- summary(colGPA.lm2)$coefficients['skipped', 't value']
df <- colGPA.lm2$df.residual
tcrit <- qt(0.01, df)
pval <- pt(tstat, df)
c(tstat, df, tcrit, pval)
```

```{r skipTest-check}
grade_this({
  if (identical(unname(.result), .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```
