---
title: "Chapter 14"
subtitle: "Advanced Panel Data Methods"
description: >
  This tutorial introduces advanced panel methods. 
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = TRUE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE, fig.height= 9, fig.width=8)
```

# Correlated Random Effects

$$ln(Crime Rate_{it}) = \beta_0 + \beta_1ln(P_{it}[Arrest]) + \beta_2ln(P_{it}[Conviction]) + \beta_3ln(P_{it}[Prison]) + $$
$$\beta_4ln(AveSentence_{it}) + \beta_5ln(PolicePC_{it}) + \beta_6West_{i} + \beta_7Central_{i} + \beta_8Urban_{i} + a_{i} + u_{it}$$

Estimate this model the following ways: 

  1. Use a standard random-effects estimator ($crmrte.re3$)
  2. Use a random-effects estimator, adding the *between transforms* ($variable.b$) of all of the time-varying variables as additional controls ($crmrte.cre1$). 
  3. Use a random-effects estimator with between-effects controls, adding region ($west$ and $central$) and $urban$ as controls ($crmrte.cre2$).
  4. Use a fixed-effects estimator, *without* between-effects controls, adding region ($west$ and $central$) and $urban$ as controls ($crmrte.fe5$).

Create a $stargazer$ text table of these estimates and the fixed-effects estimator ($crmrte.fe4$). 

```{r cre, exercise = TRUE}

```

```{r cre-hint}
crmrte.cre1 <- plm(formula, data, model, effects)
crmrte.cre2 <- plm(formula, data, model, effects)
crmrte.fe5 <- plm(formula, data, effects)
stargazer(list, type)
```

```{r cre-solution}
crmrte.cre1 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + 
                    lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.cre2 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central +
                    urban + lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.fe5 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central + 
                   urban, data = crime4.p, effects = "individual")
stargazer(crmrte.cre1, crmrte.cre2, crmrte.fe5, crmrte.fe4, type = 'text')
```

```{r cre-check}
grade_code()
```

# Other Adventures in Panel Data

## Fixed or random effects? Which should I choose?

## Fixing standard errors for heteroscedasticity/clustering

  1. "Robust" standard errors
  2. Individual-specific clustered SEs

## Other data structures

  1. Cluster sampling
  2. Multilevel models

# Clustering Standard Errors

Clustering Standard Errors: Differences in standard errors within panels

Replicate the correlated random effects model from (4) in the previous exercise using clustered standard errors (package = lmtest; function = coeftest(model, vcov.)). Perform the clustering two ways: 

  1. Using the classical OLS standard errors; 
  2. Using the heteroscedasticity-consistent standard errors (vcovHC); 
  3. Using the small-sample corrected HC standard errors (type = 'sss'). 

Summarize your results using stargazer.

```{r cluster, exercise = TRUE}

```

```{r cluster-hint}
crmrte.cre2HC0 <-coeftest(model, vcovHC)
crmrte.cre2SSS <-coeftest(model, vcovHC(crmrte.cre2, type="sss"))
stargazer(crmrte.cre2, crmrte.cre2HC0, crmrte.cre2SSS, type = 'text')
```

```{r cluster-solution}
crmrte.cre2HC0 <-coeftest(crmrte.cre2, vcovHC)
crmrte.cre2SSS <-coeftest(crmrte.cre2, vcovHC(crmrte.cre2, type="sss"))
stargazer(crmrte.cre2, crmrte.cre2HC0, crmrte.cre2SSS, type = 'text')
```

```{r cluster-check}
grade_code()
```

