---
output: slidy_presentation
runtime: shiny_prerendered
title: "Chapter 14 \n\n Advanced Panel Data Methods"
author: "James Bang"
description: >
  This tutorial introduces advanced panel methods. 
---

```{r setup, include=FALSE}
library(learnr)
library(wooldridge)
library(plm)
library(lmtest)
library(sandwich)
library(gradethis)
library(gtsummary)
library(stargazer)
library(lmtest)
gradethis::gradethis_setup()
crime4 <- wooldridge::crime4
crime4.p <- pdata.frame(crime4, index=c("county", "year") )
crime4.p$lcrmrte.b <- Between(crime4.p$lcrmrte)
crime4.p$lcrmrte.w <- Within(crime4.p$lcrmrte)
crime4.p$lprbarr.b <- Between(crime4.p$lprbarr)
crime4.p$lprbarr.w <- Within(crime4.p$lprbarr)
crime4.p$lprbconv.b <- Between(crime4.p$lprbconv)
crime4.p$lprbconv.w <- Within(crime4.p$lprbconv)
crime4.p$lprbpris.b <- Between(crime4.p$lprbpris)
crime4.p$lprbpris.w <- Within(crime4.p$lprbpris)
crime4.p$lavgsen.b <- Between(crime4.p$lavgsen)
crime4.p$lavgsen.w <- Within(crime4.p$lavgsen)
crime4.p$lpolpc.b <- Between(crime4.p$lpolpc)
crime4.p$lpolpc.w <- Within(crime4.p$lpolpc)
crmrte.fe1 <- lm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + factor(county), 
                  data = crime4.p)
crmrte.fe2 <- lm(lcrmrte.w ~ lprbarr.w + lprbconv.w + lprbpris.w + lavgsen.w + lpolpc.w, 
                  data = crime4.p)
crmrte.fe3 <- plm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
                  data = crime4.p, model = "within")
crmrte.fe4 <- plm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
                  data = crime4.p, effects = "individual")
crmrte.be1 <- lm(lcrmrte.b ~ lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                 data = crime4.p)
crmrte.re1 <- plm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.re2 <- plm(log(crmrte) ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc, 
                  data = crime4.p, effects = "none")
crmrte.cre1 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + 
                    lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.cre2 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central +
                    urban + lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.fe5 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central + 
                   urban, data = crime4.p, effects = "individual")
knitr::opts_chunk$set(echo = FALSE)
```

# Correlated Random Effects

$$ln(Crime Rate_{it}) = \beta_0 + \beta_1ln(P_{it}[Arrest]) + \beta_2ln(P_{it}[Conviction]) + \beta_3ln(P_{it}[Prison]) + $$
$$\beta_4ln(AveSentence_{it}) + \beta_5ln(PolicePC_{it}) + \beta_6West_{i} + \beta_7Central_{i} + \beta_8Urban_{i} + a_{i} + u_{it}$$

Estimate this model the following ways: 

  1. Use a standard random-effects estimator ($crmrte.re3$)
  2. Use a random-effects estimator, adding the *between transforms* ($variable.b$) of all of the time-varying variables as additional controls ($crmrte.cre1$). 
  3. Use a random-effects estimator with between-effects controls, adding region ($west$ and $central$) and $urban$ as controls ($crmrte.cre2$).
  4. Use a fixed-effects estimator, *without* between-effects controls, adding region ($west$ and $central$) and $urban$ as controls ($crmrte.fe5$).

Create a $stargazer$ text table of these estimates and the fixed-effects estimator ($crmrte.fe4$). 

```{r cre, exercise = TRUE}

```

```{r cre-hint}
crmrte.cre1 <- plm(formula, data, model, effects)
crmrte.cre2 <- plm(formula, data, model, effects)
crmrte.fe5 <- plm(formula, data, effects)
stargazer(list, type)
```

```{r cre-solution}
crmrte.cre1 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + 
                    lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.cre2 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central +
                    urban + lprbarr.b + lprbconv.b + lprbpris.b + lavgsen.b + lpolpc.b, 
                  data = crime4.p, model = "random", effects = "none")
crmrte.fe5 <- plm(lcrmrte ~ lprbarr + lprbconv + lprbpris + lavgsen + lpolpc + west + central + 
                   urban, data = crime4.p, effects = "individual")
stargazer(crmrte.cre1, crmrte.cre2, crmrte.fe5, crmrte.fe4, type = 'text')
```

```{r cre-check}
grade_code()
```

# Other Adventures in Panel Data

## Fixed or random effects? Which should I choose?

## Fixing standard errors for heteroscedasticity/clustering

  1. "Robust" standard errors
  2. Individual-specific clustered SEs

## Other data structures

  1. Cluster sampling
  2. Multilevel models

# Clustering Standard Errors

Clustering Standard Errors: Differences in standard errors within panels

Replicate the correlated random effects model from (4) in the previous exercise using clustered standard errors (package = lmtest; function = coeftest(model, vcov.)). Perform the clustering two ways: 

  1. Using the classical OLS standard errors; 
  2. Using the heteroscedasticity-consistent standard errors (vcovHC); 
  3. Using the small-sample corrected HC standard errors (type = 'sss'). 

Summarize your results using stargazer.

```{r cluster, exercise = TRUE}

```

```{r cluster-hint}
crmrte.cre2HC0 <-coeftest(model, vcovHC)
crmrte.cre2SSS <-coeftest(model, vcovHC(crmrte.cre2, type="sss"))
stargazer(crmrte.cre2, crmrte.cre2HC0, crmrte.cre2SSS, type = 'text')
```

```{r cluster-solution}
crmrte.cre2HC0 <-coeftest(crmrte.cre2, vcovHC)
crmrte.cre2SSS <-coeftest(crmrte.cre2, vcovHC(crmrte.cre2, type="sss"))
stargazer(crmrte.cre2, crmrte.cre2HC0, crmrte.cre2SSS, type = 'text')
```

```{r cluster-check}
grade_code()
```

