---
title: "Chapter 4"
subtitle: "Multiple Regression Analysis - Inference"
description: "This tutorial introduces inference for multiple-parameter tests following multiple regression."
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = FALSE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```

## Crime by Region

### Estimation

Estimate the model from the previous practice exercises *with and without region variables*. Report the output using `stargazer()`.
$$\begin{align}
ln(CrimeRate) = &\beta_0 + \beta_1ln(LawExpendPC) + \beta_2ln(IncomePC) + \beta_3ln(PopulationDensity) + \\
& \beta_4West + \beta5Northeast + \beta_6South + u
\end{align}$$  
```{r crime, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r crime-solution}
library(stargazer)
crime2 <- wooldridge::crime2
crime.lm1 <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden) + west + nrtheast + south, 
                data = crime2)
crime.lmr <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden), data = crime2)
stargazer(crime.lm1, crime.lmr, type = 'text')
```

```{r crime-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Differences between (Non-Baseline) Regions

The excluded ("baseline") is the Midwest region, which does not have a separately-defined variable in the dataset. The t-tests for the included regions (West, Northeast, and South) represent a test involving the comparison of these other regions to the Midwest region.  
Test whether the crime rates in the Northeast region differ significantly from crime rates in the South region using *both* `glht()` *and* `lht()`.  
Report your result as a `list()` consisting of (1) the output of the `glht()` test, and (2) the output of the `lht()` test.

```{r lincom, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r lincom-hint-1}
# `glht()` from the `multcomp` package requires: 
  # an `lm` or other model with estimated parameters 
  # a linear function (`linfct`) of the variables in quotation marks
  # e.g. `glht(lm, linfct = "x1 - x2 = 0")`
# The output value for `glht()` does not report the full results for the test. Use `summary.glht()` to get the test statistic, p-value, df, etc. 
```

```{r lincom-hint-2}
# `lht()` from the `car` (companion to applied regression) package requires: 
  # a model, a hypothesis matrix, `hypothesis.matrix`, representing the left-hand side 
  # a vector of right-hand side numerical values, `rhs` (same dimensions as hypothesis.matrix) 
  # e.g. `lht(lm, hypothesis.matrix = "x1 - x2", rhs = 0)` 
```

```{r lincom-setup}
crime2 <- wooldridge::crime2
crime.lm1 <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden) + west + nrtheast + south, data = crime2)
```

```{r lincom-solution}
library(multcomp)
library(car)
list(glht(crime.lm1, linfct = "nrtheast - south = 0"),
     lht(crime.lm1, hypothesis.matrix = c("nrtheast - south"), rhs = 0))
```

```{r lincom-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Jointly Testing Regional Effects 

Calculate the test statistic and p-value for the presence of risk-adjusted (controlling for other factors) regional differences in crime rates *by hand* using the restricted and unrestricted $R^2$.  
Report the result as a `c()` vector consisting of (1) the F-statistic; and (2) the p-value. 

```{r crimeF, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r crimeF-hint}
# The numerator degrees of freedom equals the number of parameter restrictions in the test. 
# You can hard-code this by just counting (west, nrtheast, and south is three). 
# You can also get this from the difference in the residual degrees of freedom stored by the `lm` objects.
  # Remember that the restricted model has more residual df than the unrestricted (fewer parameters).
  # Number of restrictions = `lm.restricted$df.residual` - `lm.unrestricted$df.residual`
```

```{r crimeF-setup}
crime2 <- wooldridge::crime2
crime.lm1 <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden) + west + nrtheast + south, data = crime2)
crime.lmr <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden), data = crime2)
```

```{r crimeF-solution}
r2.u <- summary(crime.lm1)$r.squared 
r2.r <- summary(crime.lmr)$r.squared 
c(F <- ((r2.u-r2.r)/(crime.lmr$df.residual-crime.lm1$df.residual)) / ((1-r2.u)/crime.lm1$df.residual),
  1-pf(F, crime.lmr$df.residual-crime.lm1$df.residual, crime.lm1$df.residual))
```

```{r crimeF-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

### Using `lht()`

Perform the test again using `lht()`.

```{r lht, exercise=TRUE, exercise.reveal_solution = TRUE}

```

```{r lht-setup}
crime2 <- wooldridge::crime2
crime.lm1 <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden) + west + nrtheast + south, data = crime2)
crime.lmr <- lm(log(crmrte) ~ log(lawexpc) + log(pcinc) + log(popden), data = crime2)
```

```{r lht-solution}
lht(crime.lm1, 
    hypothesis.matrix = c('west', 'nrtheast', 'south'), 
    rhs = c(0, 0, 0))
```

```{r lht-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```
