---
title: "Chapter 10" 
subtitle: "Basic Regression Analysis with Time Series Data"
description: >
  This tutorial introduces basic issues specific to time series in regression analysis.
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = TRUE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE, fig.height= 9, fig.width=8)
```

## The Nature of Time Series Data

Temporal ordering of observations; cannot be arbitrarily reordered  
Serial correlation  
Nonindependence of observations  

### Examples

1. Static Models:  
Inflation and Unemployment (Phillips Curve)  
$$Inflation_t = \beta_0 + \beta_1 Unemployment + u_t$$  
Murder Rates  
$$Murder\ Rate_t = \beta_0 + \beta_1 Conviction\ Rate_t + \beta_2 Unemployment_t + \beta_3 Male\ Youth_t + u_t$$  
2. Finite Distributed Lag Models:  
Fertility Rates  
$$Fertility\ Rate_t = \alpha_0 + \delta_0 Tax\ Exemption_t + \delta_1 Tax\ Exemption_{t-1} + \delta_2 Tax\ Exemption_{t-1} + u_t$$  
Effect of Temporary versus Permanent Changes in the Treatment  
$Impact\ Multiplier = \delta_0$
$Long\text{-}Run\ Multiplier = \delta_0 + \delta_1 + \delta_2$

## Assumptions  

### Assumptions needed for $\hat\beta_OLS$ to be "BLUE"

1.  Linear in the parameters  
2.  No perfect multicollinearity  
3.  $E(u|X)=0$ (strict exogeneity)  
    a.  No omitted variables  
    b.  No endogenous regressors  
    c.  No measurement error  
    d.  No included "colliders" (e.g. post-treatment effects)  
4.  $Var(u) = \sigma^2$ (homoskedasticity)  
5.  $X_i$'s and $Y_i$'s are independently and identically distributed
(random sampling)  
6.  $E(X^4) < 0$, $E(Y^4) < 0$ (no extreme "outliers")  
Additional assumption for $\hat\beta_{OLS}$ to be "MVUE"  
7.  $u \sim N(0, \sigma^2)$ (normal distribution of errors)  

### Strong Assumptions for Time Series Data  

Strict independence (2), exogeneity (4), and homoskedasticity (5) may be
unrealistic. Replace with:  
1.  Linear in the parameters  
2.  No perfect multicollinearity  
3.  $E(u_t|X)=0$ (strict exogeneity)  
4.  $Var(u) = \sigma^2$ (homoskedasticity)  
5.  $Corr(u_s, u_t |X) = 0, t \neq 0$ (independence - no serial correlation)  
Additional assumption for $\hat\beta_{OLS}$ to be "MVUE"  
6.  $u \sim N(0, \sigma^2)$ (normal distribution of errors)  

## Example: Static Model of the Phillips Curve

Use the `phillips` dataset in the `wooldridge` package containing data on US unemployment and inflation from 1948-2003 to estimate the following model:  
$$Inflation_t = \beta_0 + \beta_1 Unemployment_t$$
Summarize your output. 

```{r phillips, exercise = TRUE}

```

```{r phillips-solution}
phillips <- wooldridge::phillips
phillips.lm1 <- lm(inf ~ unem, data = phillips)
summary(phillips.lm1)
```

```{r phillips-check}
grade_this({
  if (identical(.result, .solution)) {
    pass(random_praise())
  }
  fail(random_encouragement())
})
```

## Example: Finite Distributed Lag Model of Fertility

Use the `fertil3` dataset in the `wooldridge` package containing data on US fertility rates and tax exemptions from 1913-1984 to estimate the following models:

1. $Fertility_t = \alpha_0 + \delta_1 Tax\ Exemption_t + \beta_1 WW2_t + \beta_2 Pill_t + u_t$ 

2. $Fertility_t = \alpha_0 + \delta_1 Tax\ Exemption_t + \delta_1 Tax\ Exemption_{t-1} + \delta_1 Tax\ Exemption_{t-1} + \beta_1 WW2_t + \beta_2 Pill_t + u_t$

## Trends and Seasonality  

### Time Trends

1.  Linear Time Trend  
$$\Delta y_t = y_t - y_{t-1} = \alpha_1 + u_t$$
$$y_t = \alpha_0 + \alpha_1 t + u_t$$

2.  Exponential Time Trend
$$\Delta log(y_t) = \frac{\Delta y_t}{y_t} = \alpha_1 + u_t$$
$$\log(y_t) = \alpha_0 + \alpha_1 t + u_t$$
3.  Nonparametric Time Trend (Panel Data Only)

$$y_{it} = \alpha_0 + \sum_{t=1}^T \alpha_t d_t + X_{it} \beta + u_t$$
Also known as time fixed effects. 

### Seasonality

Adjusting for seasonality merely involves including dummy variables for each month or quarter. 

$$y_t = \alpha_0 + \alpha_1 Q1_t + \alpha_2 Q2_t + \alpha_3 Q3_t + X_t \beta + u_t$$
