---
title: "Chapter 15"
subtitle: "Alternatives to Instrumental Variables"
author: 
  name: "Created by [Jim Bang](http://www.github.com/bangecon)"
  email: BangJamesT@sau.edu
  affiliation: "[St. Ambrose University](http://www.sau.edu)"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: 72
---

<style type="text/css">
h1{font-size: 24pt}
h2{font-size: 20pt}
h3{font-size: 18pt}
h4,h5,h6{font-size: 16pt}
body{font-size: 16pt}
</style>

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.eval = TRUE, exercise.reveal_solution = TRUE)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE, fig.height= 9, fig.width=8)
```

## Simple Mean Comparison

Question: Were first-class passengers more likely to survive on the Titanic? 

```{r simpleMeans, echo = TRUE}
titanic <- read_dta("https://raw.github.com/scunning1975/mixtape/master/titanic.dta") 
titanic$FirstClass <- (titanic$class == 1)
titanic$age <- (titanic$age == 1)
titanic$sex <- (titanic$sex == 0)
titanic$survived <- (titanic$survived == 1)
colnames(titanic) <- c("Class", "Adult", "Female", "Survived", "FirstClass")
add_p(tbl_summary(titanic[,2:5], by = FirstClass))
```

## Controlled Least-Squares Comparison

Using OLS and controlling for age (adult versus child) and gender tells a slightly different story: 

```{r controlledOLS, echo = TRUE, results = 'asis'}
titanic.lm <- lm(Survived ~ FirstClass + Adult + Female, data = titanic)
stargazer(titanic.lm, type = 'html')
```

## Matching: A Brief Overview of Methods

1. Propensity Scores
2. Distance Matching
    - Euclidian Distance
    - Mahalanobis Distance
3. Exact Matching
4. Coarsened Exact Matching

Most matching mixes a combination of these. 

Sometimes, we even use a controlled regression *after* matching on some of the observables.

## Matched Mean Comparison

Matching (exactly) according to age (adult versus child) and gender tells an even more different story: 

```{r matchedMeans, echo = TRUE}
classMatch <- matchit(FirstClass ~ Adult + Female, data = titanic, method = 'nearest')
titanic.matchit  <- match.data(classMatch)
add_p(tbl_summary(titanic.matchit[,2:5], by = FirstClass))
```

